---
title: "01-Analysis scRNAseq"
author: "Veronika Niederlova"
date: "2024-10-01"
output: rmdformats::downcute
---

<style>

div.white pre { background-color:white; color:black}
div.white pre.r { background-color:white; color:black}

</style>

<div class = "white">

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE, warning = FALSE, message = FALSE, error = TRUE, cache = TRUE, cache.lazy = FALSE)

library("Matrix")
library(Seurat)
library(pheatmap)
library(DT)
library(data.table)
library(dplyr)
library(here)
library(ggplot2)
library(SingleCellExperiment)
library(scater)
library(mclust)
library(kableExtra)
library(cowplot)
library(tidyverse)
library(reshape)
library(fgsea)
library(SingleR)
library(topGO)
library(annotate)
library("org.Mm.eg.db")
library(biomaRt)
library(ReactomePA)
library(patchwork)
library(STACAS)

convertHumanGeneList2 <- function(x){

require("biomaRt")
human = useMart("ensembl", dataset = "hsapiens_gene_ensembl", host = "https://dec2021.archive.ensembl.org/")
mouse = useMart("ensembl", dataset = "mmusculus_gene_ensembl", host = "https://dec2021.archive.ensembl.org/")

genesV2 = getLDS(attributes = c("hgnc_symbol"), filters = "hgnc_symbol", values = x , mart = human, attributesL = c("mgi_symbol"), martL = mouse, uniqueRows=T)

humanx <- unique(genesV2[, 2])

# Print the first 6 genes found to the screen
print(head(humanx))
return(humanx)
}

ggtheme <- function() {
  theme(
    axis.text = element_text(size = 20),
    axis.title = element_text(size = 20),
    text = element_text(size = 20, colour = "black"),
    legend.text = element_text(size = 20),
    legend.key.size =  unit(10, units = "points")
    
  )
}


  
  reduced_annot  <- function(seurat_object, tier_name, n_cells_to_other){
    md  <- seurat_object@meta.data
    tier_name_red  <- paste0(tier_name,"_red")
    cell_types_keep  <- (md  %>% group_by_at(tier_name)  %>% tally()  %>% arrange(desc(n)) %>% dplyr::filter(n > n_cells_to_other))[[tier_name]]
    md[[tier_name_red]]  <- if_else(md[[tier_name]] %in% cell_types_keep,as.character(md[[tier_name]]),"Other")
    seurat_object  <- AddMetaData(seurat_object, metadata = md[[tier_name_red]], tier_name_red)
    return(seurat_object)
}

  rank_score_func <- function(df){

df <- df %>% mutate(score = -1*log(p_val_adj+(10^-310))*avg_log2FC*(pct.1/(pct.2+10^-300)))

return(df)
  }
  
```

# Loading data for analysis recapitulation 

```{r}
funda_merge <- readRDS("./data/seu_merged_nod_mapping.rds")
```

# Introduction

This is analysis of the first experiment of RNA sequencing of tissues from NOD mice on gluten free diet (GFD) and mice on standard diet (STD). 

In this run, we processed:

*Well 1:*

- spleens from 5 mice on GFD, mouseIDs 1-5
- spleens from 4 mice on STD, mouseIDs 6-9

*Well 2:*
- pancreatic lymph nodes (pLN) from 5 mice on GFD, mouseIDs 1-5
- pLN from 4 mice on STD, mouseIDs 6-9

Samples were multiplexed using the TotalSeq-C anti-mouse hashtag antibodies (anti-CD45 clone 30-F11, anti-H-2 clone M1/42, Biolegend, #155869 (MH5), #155871 (MH6), #155873 (MH7), #155875 (MH8), #155877 (MH9) and #155879 (MH10)). Cells were loaded onto a 10x Chromium machine (10x Genomics) aiming at the yield of 1200 cells per sample and processed with Feature Barcode technology for Cell Surface Protein protocol (#CG000186 Rev D) with the Chromium Single Cell 5’ Library & Gel Bead and Chromium Single Cell 5' Feature Barcode Library kits (10x Genomics, #PN-1000014, #PN-1000020, #PN-1000080, #PN-1000009, #PN-1000084). Resulting cDNA libraries were sequenced on NovaSeq 6000 (Illumina) with the S1 Reagent Kit (100 or 300 cycles, Illumina, #20012865, #20012863).

The initial processing was done by Juraj Michalik: 
- Alignment of FASTQ with Cellranger tool (`cellranger-5.0.1`)
- Mapping reads to NOD genome (`Mus_musculus_nodshiltj.NOD_ShiLtJ_v1.102`)
- CellRanger outputs are attached (`W4_web_summary.html`, `W3_web_summary.html`)


# Spleen

## Load the data

Read data from well 3 - Spleen. 

```{r}

seu <- readRDS("G://48_lab/DATA_scRNAseq/Experiments_lab48/Exp11/04_Raw_data/Rds/exp11W3_init.rds")

seu[["percent.mt"]] <- PercentageFeatureSet(seu, pattern = "^mt-")
seu[["percent.rt"]] <- PercentageFeatureSet(seu, pattern = "^Rp[ls]")

# Remove TRAV, TRBV, mito and ribo genes: spleen
seu_annots <- seu@meta.data
seu_counts <- seu[!grepl('Tr[ab]v|mt-|Rp[ls]|MGP-', rownames(seu)),]
```

Filtering out ```nrow(seu)-nrow(seu_counts)``` genes:

```{r}
data.table(rownames(seu)[which(grepl('Tr[ab]v|mt-|Rp[ls]|MGP-', rownames(seu)))]) %>% 
  kbl() %>% 
  kable_styling(font_size = 11,
                bootstrap_options = c("striped", "hover", 
                                      "condensed", "responsive")) %>% 
  scroll_box(width = "100%", height = "400px")
```

## Clustering and automated annotation

```{r, fig.width=19, fig.height=5.5}

seu <- CreateSeuratObject(counts = seu_counts[["RNA"]]@counts, 
                          project = "seu", 
                          min.cells = 3, 
                          min.features = 200, 
                          meta.data = seu_annots)

  seu <- NormalizeData(object = seu)
  seu <- ScaleData(seu, verbose = FALSE)
  seu <- FindVariableFeatures(seu, selection.method = "vst", nfeatures = 1000, 
                                       verbose = FALSE)
  seu <- RunPCA(seu, npcs = 20, verbose = FALSE)
  seu <- RunUMAP(seu, reduction = "pca", dims = 1:20)
  seu <- FindNeighbors(seu, dims = 1:20)
  seu <- FindClusters(seu, resolution = 0.9)

counts <- GetAssayData(seu[["RNA"]])

seu_singler <- CreateSinglerObject(counts=counts,
  project.name= "E07_Tregs", # choose
  min.genes = 200, # ignore cells with fewer than 200 transcripts
  technology = "10x", # choose
  species = "Mouse",
  citation = "Our 10x data exp 01-05", # choose
  normalize.gene.length = FALSE,        # needed for full-length platforms (e.g. smartseq)
  variable.genes = "de",  # see vignette
  fine.tune = FALSE, # TRUE would take very long
  reduce.file.size = TRUE, # leave out less-often used fields 
  do.signatures = FALSE,
  do.main.types = TRUE,
  numCores = 16)

seu <- AddMetaData(seu,
seu_singler$singler[[1]]$SingleR.single$labels,
                                 col.name = "Immgen_annot_single")
seu <- AddMetaData(seu,
seu_singler$singler[[1]]$SingleR.single.main$labels,
                                 col.name = "Immgen_annot_single_main")
seu <- AddMetaData(seu,
seu_singler$singler[[2]]$SingleR.single.main$labels,
                                 col.name = "MouseRNAseq_single_main")
seu <- AddMetaData(seu,
seu_singler$singler[[2]]$SingleR.single$labels,
                                 col.name = "MouseRNAseq_single")

cowplot::plot_grid(DimPlot(seu, group.by = "MouseRNAseq_single_main"),
    DimPlot(seu, group.by = "Immgen_annot_single_main"), ncol = 2)

seu_spleen <- seu

DimPlot(seu_spleen, group.by = "Immgen_annot_single_main", raster = T, raster.dpi = c(900,900))
ggsave("./fig/spleen_annotations.svg", width = 16, height = 12, units = "cm")

seu_spleen_filt <- subset(seu_spleen, nFeature_RNA > 750 & percent.mt < 7.5 & seurat_clusters %in% 0:14 &
                          Immgen_annot_single_main %in% c("T cells", "NK cells", "NKT", "Tgd", "ILC")  )

seu_spleen$keep <- colnames(seu_spleen) %in% colnames(seu_spleen_filt)


DimPlot(seu_spleen, group.by = "keep", raster = T, raster.dpi = c(900,900), pt.size = 3.5, cols = c("red","grey88"))

ggsave("./fig/spleen_qc.svg", width = 13, height = 12, units = "cm")
```


## Add metadata

We need to connect the hashtags of the analyzed mice with MouseIDs and matching diets. 

```{r}
seu_spleen$hashtags %>% table

md_spleen <- seu_spleen@meta.data 
md_spleen <- md_spleen %>% mutate(MouseID = str_replace_all(string = hashtags, pattern = "_Hashtag", replacement = ""))
md_spleen <- md_spleen %>% mutate(Diet = if_else(MouseID %in% c("Mouse_1", "Mouse_2", "Mouse_3", "Mouse_4", "Mouse_5"), "GFD", "Standard"))

seu_spleen@meta.data <- md_spleen
```

```{r}
DimPlot(seu_spleen, label = T)
```


## Remove contaminating and low quality cells

```{r}
VlnPlot(seu_spleen, features = c("nCount_RNA","nFeature_RNA","percent.mt","percent.rt"), ncol = 2)
seu_spleen_filt <- subset(seu_spleen, nFeature_RNA > 750 & percent.mt < 7.5 & seurat_clusters %in% 0:14)
```



Filtering removed ```ncol(seu_spleen)-ncol(seu_spleen_filt)``` cells


```{r}
seu_spleen_filt <- NormalizeData(object = seu_spleen_filt)
  seu_spleen_filt <- ScaleData(seu_spleen_filt, verbose = FALSE)
  seu_spleen_filt <- FindVariableFeatures(seu_spleen_filt, selection.method = "vst", nfeatures = 500, 
                                       verbose = FALSE)
  seu_spleen_filt <- RunPCA(seu_spleen_filt, npcs = 12, verbose = FALSE)
  seu_spleen_filt <- RunUMAP(seu_spleen_filt, reduction = "pca", dims = 1:12)
  seu_spleen_filt <- FindNeighbors(seu_spleen_filt, dims = 1:12)
  seu_spleen_filt <- FindClusters(seu_spleen_filt, resolution = 0.7)

  DimPlot(seu_spleen_filt, group.by = "Immgen_annot_single_main", label = T, repel = T) 
  DimPlot(seu_spleen_filt, group.by = "MouseRNAseq_single_main", label = T, repel = T) 
  DimPlot(seu_spleen_filt, label = T, repel = T) 
```


```{r fig.width=16, fig.height = 9}

 seu_spleen_filt <- reduced_annot(seu_spleen_filt, "Immgen_annot_single", n_cells_to_other = 30)

DimPlot(seu_spleen_filt, group.by = "Immgen_annot_single_red", label = T, repel = T) 
  #ggsave("fig/Immgen_dimplot_spleen_filt_Immgen_single.png", width = 40, height = 20, units = "cm")
```

## Analysis of spleen clusters

Immgen Annotations

```{r fig.width=12, fig.height = 12}
seurat_meta_data <- seu_spleen_filt@meta.data

# Create grouped dataframe, calculate the frequencies of clusters
df4 <- seurat_meta_data %>% 
  group_by(MouseID, Immgen_annot_single_red) %>% 
  summarise(n = n()) %>% 
  unique() %>% 
  mutate(freq = n / sum(n)) 

# Control - all sums should be one
# df4 %>% group_by(MouseID) %>% summarise(sum = sum(freq))

# As we've lost non-grouping variables, let's join them back
md_to_join <- seurat_meta_data %>% 
  dplyr::select(MouseID, Diet) %>% 
  unique()

df4  <- left_join(df4, md_to_join)

# The final plot

df4 %>% 
  ggplot(aes(x = Diet, y = freq)) + # you can change the x to whatever variable you're interested in
  geom_boxplot(outlier.shape = NA) + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(position=position_jitter(0.2), size = 2, aes(color = Diet)) + # in aes, you can also use shape or fill (for the shapes that allow it)
  facet_wrap(~factor(Immgen_annot_single_red, labels = str_replace(levels(factor(df4$Immgen_annot_single_red)), pattern = " \\(", replacement = " \n(")), scales = "free", ncol = 5) +
  ylab("Frequency") +
  xlab("Condition") +
  theme_classic() +
  theme(strip.background = element_blank(), panel.grid = element_blank()) + 
  theme(axis.text.x = element_text(angle = 90)) + ggtheme() 

 #ggsave("fig/clusters_boxplots_spleen_filt_Immgen_single.png", width = 27, height = 35, units = "cm")
```

Seurat clusters 

```{r}
seurat_meta_data <- seu_spleen_filt@meta.data

# Create grouped dataframe, calculate the frequencies of clusters
df4 <- seurat_meta_data %>% 
  group_by(MouseID, seurat_clusters) %>% 
  summarise(n = n()) %>% 
  unique() %>% 
  mutate(freq = n / sum(n)) 

# Control - all sums should be one
# df4 %>% group_by(MouseID) %>% summarise(sum = sum(freq))

# As we've lost non-grouping variables, let's join them back
md_to_join <- seurat_meta_data %>% 
  dplyr::select(MouseID, Diet) %>% 
  unique()

df4  <- left_join(df4, md_to_join)

# The final plot

df4 %>% 
  ggplot(aes(x = Diet, y = freq)) + # you can change the x to whatever variable you're interested in
  geom_boxplot(outlier.shape = NA) + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(position=position_jitter(0.2), size = 2, aes(color = Diet)) + 
  facet_wrap(~seurat_clusters, scales = "free", ncol = 5) +
  ylab("Frequency") +
  xlab("Condition") +
  theme_classic() +
  theme(strip.background = element_blank(), panel.grid = element_blank()) + 
  theme(axis.text.x = element_text(angle = 90))

 #ggsave("fig/clusters_boxplots_spleen_filt_clusters.png", width = 16, height = 16, units = "cm")
```



## Filter lymphoid populations

## Filter high quality & lymphoid populations

```{r}
seu_spleen_filt_lymph <- subset(seu_spleen, nFeature_RNA > 750 & percent.mt < 7.5 & Immgen_annot_single_main %in% c("T cells", "NK cells", "NKT", "Tgd", "ILC"))

seu_spleen_filt_lymph <- NormalizeData(object = seu_spleen_filt_lymph)
  seu_spleen_filt_lymph <- ScaleData(seu_spleen_filt_lymph, verbose = FALSE)
  seu_spleen_filt_lymph <- FindVariableFeatures(seu_spleen_filt_lymph, selection.method = "vst", nfeatures = 500, 
                                       verbose = FALSE)
  seu_spleen_filt_lymph <- RunPCA(seu_spleen_filt_lymph, npcs = 12, verbose = FALSE)
  seu_spleen_filt_lymph <- RunUMAP(seu_spleen_filt_lymph, reduction = "pca", dims = 1:12)
  seu_spleen_filt_lymph <- FindNeighbors(seu_spleen_filt_lymph, dims = 1:12)
  seu_spleen_filt_lymph <- FindClusters(seu_spleen_filt_lymph, resolution = 0.8)

  DimPlot(seu_spleen_filt_lymph, group.by = "Immgen_annot_single_main", label = T, repel = T) 
  DimPlot(seu_spleen_filt_lymph, group.by = "MouseRNAseq_single_main", label = T, repel = T) 
  
  DimPlot(seu_spleen_filt_lymph, label = T, repel = T) 
  
```

## Analysis of spleen clusters

Annotations by diet

```{r}

seurat_meta_data <- seu_spleen_filt_lymph@meta.data

# Create grouped dataframe, calculate the frequencies of clusters
df4 <- seurat_meta_data %>% 
  group_by(MouseID, seurat_clusters) %>% 
  summarise(n = n()) %>% 
  unique() %>% 
  mutate(freq = n / sum(n)) 

# Control - all sums should be one
# df4 %>% group_by(MouseID) %>% summarise(sum = sum(freq))

# As we've lost non-grouping variables, let's join them back
md_to_join <- seurat_meta_data %>% 
  dplyr::select(MouseID, Diet) %>% 
  unique()

df4  <- left_join(df4, md_to_join)

# The final plot

df4 %>% 
  ggplot(aes(x = Diet, y = freq)) + # you can change the x to whatever variable you're interested in
  geom_boxplot(outlier.shape = NA) + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(position=position_jitter(0.2), size = 2, aes(color = Diet)) + # in aes, you can also use shape or fill (for the shapes that allow it)
  facet_wrap(~seurat_clusters, scales = "free", ncol = 5) +
  ylab("Frequency") +
  xlab("Condition") +
  theme_classic() +
  theme(strip.background = element_blank(), panel.grid = element_blank()) + 
  theme(axis.text.x = element_text(angle = 90))

#ggsave("fig/clusters_boxplots_spleen_filt_lymph_clusters.png", width = 20, height = 15, units = "cm")
```

## Markers spleen

```{r eval=FALSE, include=TRUE}
markers_spleen <- FindAllMarkers(seu_spleen_filt_lymph, logfc.threshold = log(2))

markers_spleen %>% 
  kbl() %>% 
  kable_styling(font_size = 11,
                bootstrap_options = c("striped", "hover", 
                                      "condensed", "responsive")) %>% 
  scroll_box(width = "100%", height = "400px")
write.csv(markers_spleen, "markers_spleen.csv", row.names = T)
```



## Differentially expressed genes cluster 0 and cluster 3

```{r eval=FALSE, include=TRUE}
spl_cl1_vs_cl2 <- FindMarkers(seu_spleen_filt_lymph, ident.1 = 1, ident.2 = 2, only.pos = F)


data.table(spl_cl1_vs_cl2) %>% 
  kbl() %>% 
  kable_styling(font_size = 11,
                bootstrap_options = c("striped", "hover", 
                                      "condensed", "responsive")) %>% 
  scroll_box(width = "100%", height = "400px")

write.csv(spl_cl1_vs_cl2, "spl_cl1_vs_cl2.csv", row.names = T)

```


```{r fig.width=18, fig.height=12}

FeaturePlot(seu_spleen_filt_lymph, features = c("Cd3d","Cd4","Cd8a","Sell","Ccl5",
                                       "Foxp3","Il2ra","Pdcd1", "Cxcr5","Klrd1","Klrk1",
                                       "Isg15","Srm","Klrg1", "Mki67"), ncol = 5)
#ggsave("fig/markers1_seu_spleen_filt_lymph.png", width = 20, height = 11)

FeaturePlot(seu_spleen_filt_lymph, features = c("Ifng","Tbx21","Eomes","Bcl6","Ccr8",
                                       "Ccr10","Cxcr3","Tox", "Lag3","Ikzf2","Id2",
                                       "Id3","Prdm1","Itgae", "Cd69"), ncol = 5)
#ggsave("fig/markers2_seu_spleen_filt_lymph.png", width = 20, height = 11)

FeaturePlot(seu_spleen_filt_lymph, features = c("Gata3","Rorc","Il23r","Cxcr6","Cd27",
                                       "Entpd1","Izumo1r","Tnfrsf9", "Tnfrsf18","Tnfrsf14","Il17a",
                                       "Gzma","Gzmb","Gzmk", "Gzmm"), ncol = 5)
#ggsave("fig/markers3_seu_spleen_filt_lymph.png", width = 20, height = 11)


```

## Differential expression

Differentially expressed genes between GFD and STD.

By DeSeq spleen

```{r fig.width=20, fig.height=30}
sum_exp <- AggregateExpression(seu_spleen_filt_lymph, group.by = "MouseID", slot = "counts")

md <- data.frame(Diet = c("GF", "GF", "GF", "GF", "GF", "STD", "STD", "STD", "STD"))

# Create a DESeq dataset
library(DESeq2)
dds <- DESeqDataSetFromMatrix(countData = sum_exp$RNA,
                              colData = md,
                              design= ~ Diet) 

rownames(dds) <- rownames(sum_exp$RNA)

# Run the DESeq2 algorithm, which will give us the differentially expressed genes
dds <- DESeq(dds)

```


```{r fig.width=4, fig.height=4}
## PCA
pca = vst(dds, blind = FALSE)
pca_plot <- plotPCA(pca, intgroup = c("Diet"))
pca_plot
```


```{r}
## Results data frame
res <- results(dds)
results <- as.data.frame(dplyr::mutate(as.data.frame(res), sig=ifelse(res$padj<0.05, "FDR<0.05", "Not Sig")), row.names=rownames(res))
```

Genes downregulated in STD

```{r}
results %>% filter(log2FoldChange < 0) %>%  arrange(pvalue) %>% 
  kbl() %>% 
  kable_styling(font_size = 11,
                bootstrap_options = c("striped", "hover", 
                                      "condensed", "responsive")) %>% 
  scroll_box(width = "100%", height = "400px")
```

Genes upregulated in GFD

```{r}
results %>% filter(log2FoldChange > 0) %>%  arrange(pvalue) %>% 
  kbl() %>% 
  kable_styling(font_size = 11,
                bootstrap_options = c("striped", "hover", 
                                      "condensed", "responsive")) %>% 
  scroll_box(width = "100%", height = "400px")

plotCounts(dds, gene="Slc16a5", intgroup="Diet")
plotCounts(dds, gene="Itga4", intgroup="Diet")
plotCounts(dds, gene="Stat3", intgroup="Diet")
plotCounts(dds, gene="Il6st", intgroup="Diet")
plotCounts(dds, gene="Adam17", intgroup="Diet")
plotCounts(dds, gene="Klrd1", intgroup="Diet")
plotCounts(dds, gene="Il18r1", intgroup="Diet")

plotCounts(dds, gene="Stat", intgroup="Diet")
plotCounts(dds, gene="Ccl4", intgroup="Diet")


```

# Pancreatic lymph node

Read data from well 4 - pLN

```{r}
seu <- readRDS("G://48_lab/DATA_scRNAseq/Experiments_lab48/Exp11/04_Raw_data/Rds/exp11W4_init.rds")

seu[["percent.mt"]] <- PercentageFeatureSet(seu, pattern = "^mt-")
seu[["percent.rt"]] <- PercentageFeatureSet(seu, pattern = "^Rp[ls]")

# Remove TRAV, TRBV genes: blood young B6
seu_annots <- seu@meta.data
seu_counts <- seu[!(grepl('Trav', rownames(seu)) |
          grepl('Trbv', rownames(seu)) |
          grepl('mt-', rownames(seu)) |
            grepl('Rpl', rownames(seu)) |
          grepl('Rps', rownames(seu)) |
          grepl('MGP-', rownames(seu)) 
            ),]

seu <- CreateSeuratObject(counts = seu_counts[["RNA"]]@counts, project = "seu", min.cells = 3, min.features = 200, meta.data = seu_annots)
 seu <- NormalizeData(object = seu)
  seu <- ScaleData(seu, verbose = FALSE)
  seu <- FindVariableFeatures(seu, selection.method = "vst", nfeatures = 1000, 
                                       verbose = FALSE)
  seu <- RunPCA(seu, npcs = 20, verbose = FALSE)
  seu <- RunUMAP(seu, reduction = "pca", dims = 1:20)
  seu <- FindNeighbors(seu, dims = 1:20)
  seu <- FindClusters(seu, resolution = 0.9)

 counts <- GetAssayData(seu[["RNA"]])

seu_singler <- CreateSinglerObject(counts=counts,
  project.name= "E07_Tregs", # choose
  min.genes = 200, # ignore cells with fewer than 200 transcripts
  technology = "10x", # choose
  species = "Mouse",
  citation = "Our 10x data exp 01-05", # choose
  normalize.gene.length = FALSE,        # needed for full-length platforms (e.g. smartseq)
  variable.genes = "de",  # see vignette
  fine.tune = FALSE, # TRUE would take very long
  reduce.file.size = TRUE, # leave out less-often used fields 
  do.signatures = FALSE,
  do.main.types = TRUE,
  numCores = 16)


seu <- AddMetaData(seu,
seu_singler$singler[[1]]$SingleR.single$labels,
                                 col.name = "Immgen_annot_single")
seu <- AddMetaData(seu,
seu_singler$singler[[1]]$SingleR.single.main$labels,
                                 col.name = "Immgen_annot_single_main")
seu <- AddMetaData(seu,
seu_singler$singler[[2]]$SingleR.single.main$labels,
                                 col.name = "MouseRNAseq_single_main")
seu <- AddMetaData(seu,
seu_singler$singler[[2]]$SingleR.single$labels,
                                 col.name = "MouseRNAseq_single")


DimPlot(seu, group.by = "Immgen_annot_single_main", label = T)

seu_pLN <- seu



DimPlot(seu, group.by = "Immgen_annot_single_main", raster = T, pt.size = 4, raster.dpi = c(900,900)) + ggtheme()
ggsave("./fig/pLN_annotations.svg", width = 16, height = 12, units = "cm")

seu_pLN$keep <- colnames(seu_pLN) %in% colnames(seu_pLN_filt)

seu_pLN_filt <- subset(seu_pLN, nFeature_RNA > 750 & percent.mt < 7.5 & Immgen_annot_single_main %in% c("T cells", "NK cells", "NKT", "Tgd", "ILC"))



DimPlot(seu_pLN, group.by = "keep", raster = T, pt.size = 3, , cols = c("red2","grey88"),
        raster.dpi = c(900,900)) + ggtheme()
ggsave("./fig/pLN_qc.svg", width = 16, height = 12, units = "cm")

```


## Add mouse information

```{r}
md_pLN <- seu_pLN@meta.data 
md_pLN <- md_pLN %>% mutate(MouseID = str_replace_all(string = hashtags, pattern = "_Hashtag", replacement = ""))
md_pLN <- md_pLN %>% mutate(Diet = if_else(MouseID %in% c("Mouse_1", "Mouse_2", "Mouse_3", "Mouse_4", "Mouse_5"), "GFD", "Standard"))

seu_pLN@meta.data <- md_pLN

DimPlot(seu_pLN, group.by = "MouseID")

VlnPlot(seu_pLN, features = c("nCount_RNA","nFeature_RNA","percent.mt","percent.rt"), ncol = 2)

DimPlot(seu_pLN, group.by = "Immgen_annot_single_main", label = T, repel = T) 
DimPlot(seu_pLN, group.by = "MouseRNAseq_single_main", label = T, repel = T)

```

## Remove contaminating and low quality cells

```{r}

seu_pLN_filt <- subset(seu_pLN, nFeature_RNA > 750 & percent.mt < 7.5 & Immgen_annot_single_main %in% c("T cells", "NK cells", "NKT", "Tgd", "ILC"))

seu_pLN_filt <- NormalizeData(object = seu_pLN_filt)
  seu_pLN_filt <- ScaleData(seu_pLN_filt, verbose = FALSE)
  seu_pLN_filt <- FindVariableFeatures(seu_pLN_filt, selection.method = "vst", nfeatures = 500, 
                                       verbose = FALSE)
  seu_pLN_filt <- RunPCA(seu_pLN_filt, npcs = 12, verbose = FALSE)
  seu_pLN_filt <- RunUMAP(seu_pLN_filt, reduction = "pca", dims = 1:12)
  seu_pLN_filt <- FindNeighbors(seu_pLN_filt, dims = 1:12)
  seu_pLN_filt <- FindClusters(seu_pLN_filt, resolution = 0.7)

  DimPlot(seu_pLN_filt, group.by = "Immgen_annot_single_main", label = T, repel = T) 
  DimPlot(seu_pLN_filt, group.by = "MouseRNAseq_single_main", label = T, repel = T) 
  
  DimPlot(seu_pLN_filt, label = T, repel = T) 
  
  DimPlot(seu_pLN_filt, group.by = "Diet", label = T, repel = T) 
  
```

## Markers pLN

```{r eval=FALSE, include=TRUE}
markers_pln <- FindAllMarkers(seu_pLN_filt, logfc.threshold = log(2))
write.csv(markers_pln, "markers_pln.csv", row.names = F)

data.table(markers_pln %>% rank_score_func %>% filter(avg_log2FC > 0) %>% group_by(cluster) %>% slice_max(order_by = score, n = 20)) %>% 
  kbl() %>% 
  kable_styling(font_size = 11,
                bootstrap_options = c("striped", "hover", 
                                      "condensed", "responsive")) %>% 
  scroll_box(width = "100%", height = "400px")
```

## Differentially expressed genes cluster 0 and cluster 3

```{r eval=FALSE, include=TRUE}
pLN_cl0_vs_cl3 <- FindMarkers(seu_pLN_filt, ident.1 = 0, ident.2 = 3, only.pos = F)

pLN_cl0_vs_cl3

write.csv(pLN_cl0_vs_cl3, "pLN_cl0_vs_cl3.csv", row.names = T)

data.table(pLN_cl0_vs_cl3) %>% 
  kbl() %>% 
  kable_styling(font_size = 11,
                bootstrap_options = c("striped", "hover", 
                                      "condensed", "responsive")) %>% 
  scroll_box(width = "100%", height = "400px")
```

## Differential expression

Differentially expressed genes between GFD and STD.

By DeSeq spleen

```{r fig.width=10, fig.height=12}
sum_exp <- AggregateExpression(seu_spleen_filt_lymph, group.by = "MouseID", slot = "counts")

md <- data.frame(Diet = c("GF", "GF", "GF", "GF", "GF", "STD", "STD", "STD", "STD"))

# Create a DESeq dataset
library(DESeq2)
dds <- DESeqDataSetFromMatrix(countData = sum_exp$RNA,
                              colData = md,
                              design= ~ Diet) 

rownames(dds) <- rownames(sum_exp$RNA)

# Run the DESeq2 algorithm, which will give us the differentially expressed genes
dds <- DESeq(dds)

```


```{r}
## PCA
pca = vst(dds, blind = FALSE)
pca_plot <- plotPCA(pca, intgroup = c("Diet"))
pca_plot
```


```{r}
## Results data frame
res <- results(dds)
results <- as.data.frame(dplyr::mutate(as.data.frame(res), sig=ifelse(res$padj<0.05, "FDR<0.05", "Not Sig")), row.names=rownames(res))
```

Genes downregulated in STD

```{r}
p1 <- plotCounts(dds, gene="Mcpt8", intgroup="Diet")
p2 <- plotCounts(dds, gene="Isg15", intgroup="Diet")
p3 <- plotCounts(dds, gene="Gzmk", intgroup="Diet")
p4 <- plotCounts(dds, gene="Ifng", intgroup="Diet")
p5 <- plotCounts(dds, gene="Klrg1", intgroup="Diet")
p6 <- plotCounts(dds, gene="Cd44", intgroup="Diet")
p7 <- plotCounts(dds, gene="Klrk1", intgroup="Diet")
p8 <- plotCounts(dds, gene="Stat3", intgroup="Diet")
```


```{r}
results %>% filter(log2FoldChange > 1) %>%  arrange(pvalue) %>% 
  kbl() %>% 
  kable_styling(font_size = 11,
                bootstrap_options = c("striped", "hover", 
                                      "condensed", "responsive")) %>% 
  scroll_box(width = "100%", height = "400px")
```

Genes upregulated in GFD

```{r}
results %>% arrange(log2FoldChange) %>% 
  kbl() %>% 
  kable_styling(font_size = 11,
                bootstrap_options = c("striped", "hover", 
                                      "condensed", "responsive")) %>% 
  scroll_box(width = "100%", height = "400px")



```


## Boxplot

```{r}
seurat_meta_data <- seu_pLN_filt@meta.data

# Create grouped dataframe, calculate the frequencies of clusters
df4 <- seurat_meta_data %>% 
  group_by(MouseID, seurat_clusters) %>% 
  summarise(n = n()) %>% 
  unique() %>% 
  mutate(freq = n / sum(n)) 

# Control - all sums should be one
# df4 %>% group_by(MouseID) %>% summarise(sum = sum(freq))

# As we've lost non-grouping variables, let's join them back
md_to_join <- seurat_meta_data %>% 
  dplyr::select(MouseID, Diet) %>% 
  unique()

df4  <- left_join(df4, md_to_join)

# The final plot

df4 %>% 
  ggplot(aes(x = Diet, y = freq)) + # you can change the x to whatever variable you're interested in
  geom_boxplot(outlier.shape = NA) + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(position=position_jitter(0.2), size = 2, aes(color = Diet)) + 
  facet_wrap(~seurat_clusters, scales = "free", ncol = 5) +
  ylab("Frequency") +
  xlab("Condition") +
  theme_classic() +
  theme(strip.background = element_blank(), panel.grid = element_blank()) + 
  theme(axis.text.x = element_text(angle = 90))

 #ggsave("fig/clusters_boxplots_pLN_filt_clusters.png", width = 18, height = 10, units = "cm")
```

## Annotations

```{r fig.width=18, fig.height=12}

FeaturePlot(seu_pLN_filt, features = c("Cd3d","Cd4","Cd8a","Sell","Ccl5",
                                       "Foxp3","Il2ra","Pdcd1", "Cxcr5","Klrd1","Klrk1",
                                       "Isg15","Srm","Klrg1", "Mki67"), ncol = 5)


FeaturePlot(seu_pLN_filt, features = c("Ifng","Tbx21","Eomes","Bcl6","Ccr8",
                                       "Ccr10","Cxcr3","Tox", "Lag3","Ikzf2","Id2",
                                       "Id3","Prdm1","Itgae", "Cd69"), ncol = 5)

```

)
# Merged spleen & pLN



```{r}

seu_spleen_filt_lymph$Tissue <- "Spleen"
seu_pLN_filt$Tissue <- "pLN"

 funda_merge <- merge(seu_spleen_filt_lymph, seu_pLN_filt)
 funda_merge <- NormalizeData(object = funda_merge)
  funda_merge <- ScaleData(funda_merge, verbose = FALSE)
  funda_merge <- FindVariableFeatures(funda_merge, selection.method = "vst", nfeatures = 1000, 
                                       verbose = FALSE)
  funda_merge <- RunPCA(funda_merge, npcs = 10, verbose = FALSE)
  funda_merge <- RunUMAP(funda_merge, reduction = "pca", dims = 1:10)
  funda_merge <- FindNeighbors(funda_merge, dims = 1:10)
  funda_merge <- FindClusters(funda_merge, resolution = 1.1)
  DimPlot(funda_merge, label = T)
  

  DimPlot(funda_merge, group.by = "Tissue") 
  DimPlot(funda_merge, group.by = "Diet") 
  
  

```

```{r}
funda_merge <- readRDS("./data/seu_merged_nod_mapping.rds")
```

```{r}
DimPlot(funda_merge, group.by = "Immgen_annot_single_main", label = T, repel = T) 
DimPlot(funda_merge, group.by = "MouseRNAseq_single_main", label = T, repel = T) 
  
DimPlot(funda_merge, label = T, repel = T) 
  

  
funda_merge <- reduced_annot(funda_merge, "Immgen_annot_single", n_cells_to_other = 30)
```


```{r fig.width=20, fig.height=14}
DimPlot(funda_merge, group.by = "Immgen_annot_single_red", label = T, repel = T) 
 
```

## Manual annotations of clusters

```{r}
DimPlot(funda_merge, label = T)
```

## Cluster markers

```{r eval=FALSE, include=TRUE}
mrk <- FindAllMarkers(funda_merge, logfc.threshold = log(1.5), min.diff.pct = 0.05)
mrk <- rank_score_func(mrk) %>% arrange(cluster, desc(score))

# write.csv(mrk, "funda_merge_clusters.csv")

# Top 20 positive markers for each cluster
data.table(mrk %>% group_by(cluster) %>% filter(avg_log2FC > 0) %>% slice_max(order_by = score, n = 20)) %>% 
  kbl() %>% 
  kable_styling(font_size = 11,
                bootstrap_options = c("striped", "hover", 
                                      "condensed", "responsive")) %>% 
  scroll_box(width = "100%", height = "400px")
```

```{r}
md <- funda_merge@meta.data

md <- md %>% mutate(annotations_fine = recode(seurat_clusters, 
                                                     "0" = "00 CD4 Naive 1",
                                                     "1" = "01 CD8 Naive 1",
                                                     "2" = "02 CD4 Naive 2",
                                                     "3" = "03 CD4 Naive 3",
                                                     "4" = "04 CD8 Naive 2",
                                                     "5" = "05 CD4 Naive 4",
                                                     "6" = "06 CD4 Naive 5",
                                                     "7" = "07 CD4 Treg 1",
                                                     "8" = "08 CD4 Treg 2",
                                                     "9" = "09 Tgd and NKT ",
                                                     "10" = "10 NK cells",
                                                     "11" = "11 CD4 Tfh",
                                                     "12" = "12 CD4 Egr Myb",
                                                     "13" = "13 CD8 ISAGhi",
                                                     "14" = "14 CD4 Early act. 41BB CD69",
                                                     "15" = "15 CD8 AIMT",
                                                     "16" = "16 Proliferating",
                                                     "17" = "17 ILC",
                                                     "18" = "18 CD8 Egr"))

funda_merge@meta.data <- md
rownames(funda_merge@meta.data) <- colnames(funda_merge)
```


```{r}


DimPlot(funda_merge, group.by = "seurat_clusters", label = T, 
        raster = T, pt.size = 15, raster.dpi = c(5000,5000)) + scale_color_manual(values = c(
  "#b9ddf1", 
  "#b3e0a6",
  "#89b8da", 
  "#6a9bc3", 
  "#75bc69",
  "#4878a6" ,
  
  "#92c0df",
  "#F28E2B" ,
  "#FFBE7D",
  "#9D7660" ,
  "#B07AA1" ,
  
  "#F1CE63",
  "#499894"  ,
  "#398949",
  "#86BCB6" ,
  "#24693d",
  
  "indianred2",
  "#D4A6C8",
  "slategray"),
  labels = levels(md$annotations_fine))
ggsave(filename = "fig/dimPlot_19clusters.png", width = 20, height = 13, units = "cm")

ggsave(filename = "fig/dimPlot_19clusters.svg", width = 20, height = 13, units = "cm")



```




```{r}
# saveRDS(seu_pLN_filt, "data/seu_pLN_filt_nod_mapping.rds")
# saveRDS(seu_spleen_filt, "data/seu_spleen_filt_nod_mapping.rds")
# saveRDS(funda_merge, "data/seu_merged_nod_mapping.rds")
```

```{r eval=FALSE, include=TRUE}
#funda_merge <- readRDS("data/seu_merged_nod_mapping.rds")
#seu_pLN_filt <- readRDS("data/seu_pLN_filt_nod_mapping.rds")
#seu_spleen_filt <- readRDS("data/seu_spleen_filt_nod_mapping.rds")
```

```{r}
funda_merge <- BuildClusterTree(
  funda_merge,
  dims = 1:10,
  reorder = FALSE,
  reorder.numeric = FALSE
)

tree <- funda_merge@tools$BuildClusterTree
tree$tip.label <- levels(funda_merge$annotations_fine)

p <- ggtree::ggtree(tree, aes(x, y)) +
  scale_y_reverse() +
  ggtree::geom_tree() +
  ggtree::theme_tree() +
  ggtree::geom_tiplab(offset = 1) +
  ggtree::geom_tippoint(color = 
                          c(
  "#b9ddf1", 
  "#b3e0a6",
  "#89b8da", 
  "#6a9bc3", 
  "#75bc69",
  "#4878a6" ,
  
  "#92c0df",
  "#F28E2B" ,
  "#FFBE7D",
  "#9D7660" ,
  "#B07AA1" ,
  
  "#F1CE63",
  "#499894"  ,
  "#398949",
  "#86BCB6" ,
  "#24693d",
  
  "indianred2",
  "#D4A6C8",
  "slategray"), shape = 16, size = 5) +
  coord_cartesian(clip = 'off') +
  theme(plot.margin = unit(c(0,2.5,0,0), 'cm'))

p
ggsave('plots/cluster_tree.png', p, height = 4, width = 4)
ggsave('plots/cluster_tree.svg', p, height = 4, width = 4)
```


## Annotations coarse

```{r}
md <- funda_merge@meta.data
md <- md %>% mutate(annotations_coarse = recode(seurat_clusters, 
                                                     "0" = "CD4 Naive",
                                                     "1" = "CD8 Naive",
                                                     "2" = "CD4 Naive",
                                                     "3" = "CD4 Naive",
                                                     "4" = "CD8 Naive",
                                                     "5" = "CD4 Naive",
                                                     "6" = "CD4 Naive",
                                                     "7" = "CD4 Treg",
                                                     "8" = "CD4 Treg",
                                                     "9" = "Tgd and NKT",
                                                     "10" = "NK cells",
                                                     "11" = "CD4 Treg",
                                                     "12" = "CD4 Early activated",
                                                     "13" = "CD8 Early activated",
                                                     "14" = "CD4 Early activated",
                                                     "15" = "CD8 AIMT",
                                                     "16" = "Proliferating",
                                                     "17" = "ILC",
                                                     "18" = "CD8 Early activated"))

funda_merge@meta.data <- md
rownames(funda_merge@meta.data) <- colnames(funda_merge)


DimPlot(funda_merge, group.by = "annotations_coarse", label = T, 
        raster = T, pt.size = 15, raster.dpi = c(5000,5000)) + scale_color_manual(values = c(
  "#6a9bc3", 
  "#75bc69",
  "#F28E2B" ,
  "#9D7660" ,
  "#B07AA1" ,
  "#b9ddf1", 
  "#b3e0a6",
  "#24693d",
  
  "indianred2",
  "#D4A6C8"))


#ggsave(filename = "fig/dimPlot_10clusters.png", width = 20, height = 13, units = "cm")
#ggsave(filename = "fig/dimPlot_10clusters.svg", width = 20, height = 13, units = "cm")

```


## Cluster abundance boxplot

```{r fig.width=15, fig.height=16}
funda_merge_spl <- subset(funda_merge, Tissue == "Spleen")
```


```{r fig.width=15, fig.height=16}
seurat_meta_data <- funda_merge_spl@meta.data


# Create grouped dataframe, calculate the frequencies of clusters
df4 <- seurat_meta_data %>% 
  mutate(MouseID_tissue = paste(MouseID, Tissue)) %>% 
  group_by(MouseID_tissue, annotations_fine) %>% 
  summarise(n = n()) %>% 
  unique() %>% 
  mutate(freq = n / sum(n)) 

df4 <- df4 %>% dplyr::select(-n)  %>% 
ungroup   %>% 
pivot_wider(names_from = "annotations_fine", values_from = "freq", values_fill = 0) 
df4  <- left_join((nod.combined_filt@meta.data %>% dplyr::select(mouse_id_new) %>% unique), df3)
df4[is.na(df4)] <- 0
df4  <- df4  %>% pivot_longer(!mouse_id_new, values_to = "freq", names_to = "annotations_fine")


# Control - all sums should be one
# df4 %>% group_by(MouseID_tissue) %>% summarise(sum = sum(freq))

# As we've lost non-grouping variables, let's join them back
md_to_join <- seurat_meta_data %>% 
  dplyr::select(MouseID, Diet, Tissue) %>% 
  unique() %>% 
  mutate(MouseID_tissue = paste(MouseID, Tissue))
df4  <- left_join(df4, md_to_join)

# The final plot
df4 %>% 
  ggplot(aes(x = factor(Diet, labels = c("GFD","STD")), y = freq*100)) + # you can change the x to whatever variable you're interested in
  geom_boxplot(outlier.shape = NA, aes(fill = Diet), alpha = 0.5) +
 geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(shape=21, size = 3, position=position_jitter(width = 0.1, height = 0), 
              aes(fill = Diet)) + theme_classic() + xlab("Group") +  
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5), 
        strip.background = element_blank(),
        axis.text = element_text(color = "black", size = 16),
        axis.ticks.x = element_blank(),
        strip.text = element_text(size = 16)) + 
  scale_color_manual(values = c("dodgerblue","indianred2")) + 
  scale_fill_manual(values = c("dodgerblue","indianred2")) +
  facet_wrap(~annotations_fine, scales = "free") +
  ylim(0,NA) +
  ylab("Frequency") +
  xlab("Condition") +
  ggpubr::stat_compare_means(label.x = 1.2, label.y.npc = "top", size = 5, vjust = 0.3, label = "p.format") 

 ggsave("fig/clusters_boxplots_SPL.png", width = 25, height = 25, units = "cm", limitsize = FALSE)
ggsave("fig/clusters_boxplots_SPL.svg", width = 25, height = 25, units = "cm", limitsize = FALSE)

```


```{r fig.width=15, fig.height=16}
funda_merge_pln <- subset(funda_merge, Tissue == "pLN")
```


```{r fig.width=15, fig.height=16}
seurat_meta_data <- funda_merge_pln@meta.data


# Create grouped dataframe, calculate the frequencies of clusters
df4 <- seurat_meta_data %>% 
  mutate(MouseID_tissue = paste(MouseID, Tissue)) %>% 
  group_by(MouseID_tissue, annotations_fine) %>% 
  summarise(n = n()) %>% 
  unique() %>% 
  mutate(freq = n / sum(n)) 

# Control - all sums should be one
# df4 %>% group_by(MouseID_tissue) %>% summarise(sum = sum(freq))

# As we've lost non-grouping variables, let's join them back
md_to_join <- seurat_meta_data %>% 
  dplyr::select(MouseID, Diet, Tissue) %>% 
  unique() %>% 
  mutate(MouseID_tissue = paste(MouseID, Tissue))
df4  <- left_join(df4, md_to_join)

# The final plot
df4 %>% 
  ggplot(aes(x = factor(Diet, labels = c("GFD","STD")), y = freq*100)) + # you can change the x to whatever variable you're interested in
  geom_boxplot(outlier.shape = NA, aes(fill = Diet), alpha = 0.5) +
 geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(shape=21, size = 3, position=position_jitter(width = 0.1, height = 0), 
              aes(fill = Diet)) + theme_classic() + xlab("Group") +  
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5), 
        strip.background = element_blank(),
        axis.text = element_text(color = "black", size = 16),
        axis.ticks.x = element_blank(),
        strip.text = element_text(size = 16)) + 
  scale_color_manual(values = c("dodgerblue","indianred2")) + 
  scale_fill_manual(values = c("dodgerblue","indianred2")) +
  facet_wrap(~annotations_fine, scales = "free") +
  ylim(0,NA) +
  ylab("Frequency") +
  xlab("Condition") +
  ggpubr::stat_compare_means(label.x = 1.2, label.y.npc = "top", size = 5, vjust = 0.3, label = "p.format") 

 ggsave("fig/clusters_boxplots_pLN.png", width = 25, height = 25, units = "cm", limitsize = FALSE)
ggsave("fig/clusters_boxplots_pLN.svg", width = 25, height = 25, units = "cm", limitsize = FALSE)


```


## Boxplots annotations coarse 

```{r fig.width=15, fig.height=16}

seurat_meta_data <- funda_merge@meta.data

# Create grouped dataframe, calculate the frequencies of clusters
df4 <- seurat_meta_data %>% 
  mutate(MouseID_tissue = paste(MouseID, Tissue)) %>% 
  group_by(MouseID_tissue, annotations_coarse) %>% 
  summarise(n = n()) %>% 
  unique() %>% 
  mutate(freq = n / sum(n)) 

# Control - all sums should be one
# df4 %>% group_by(MouseID_tissue) %>% summarise(sum = sum(freq))

# As we've lost non-grouping variables, let's join them back
md_to_join <- seurat_meta_data %>% 
  dplyr::select(MouseID, Diet, Tissue) %>% 
  unique() %>% 
  mutate(MouseID_tissue = paste(MouseID, Tissue))
df4  <- left_join(df4, md_to_join)

# The final plot
df4 %>% 
  mutate(Tissue_Diet = paste(Tissue, Diet)) %>% 
  ggplot(aes(x = Tissue_Diet, y = freq)) + # you can change the x to whatever variable you're interested in
  geom_boxplot(outlier.shape = NA) + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(position=position_jitter(0.2), size = 2, aes(color = Diet)) + # in aes, you can also use shape or fill (for the shapes that allow it)
  facet_wrap(~annotations_coarse, scales = "free") +
  ylab("Frequency") +
  xlab("Condition") +
  theme_classic() +
  theme(strip.background = element_blank(), panel.grid = element_blank()) + 
  theme(axis.text.x = element_text(angle = 90))

# #ggsave("fig/clusters_boxplots.png", width = 25, height = 25, units = "cm")

```

```{r}

seurat_meta_data <- funda_merge_spl@meta.data


# Create grouped dataframe, calculate the frequencies of clusters
df4 <- seurat_meta_data %>% 
  mutate(MouseID_tissue = paste(MouseID, Tissue)) %>% 
  group_by(MouseID_tissue, annotations_coarse) %>% 
  summarise(n = n()) %>% 
  unique() %>% 
  mutate(freq = n / sum(n)) 

df4 <- df4 %>% dplyr::select(-n)  %>% 
ungroup   %>% 
pivot_wider(names_from = "annotations_coarse", values_from = "freq", values_fill = 0) 
df4  <- left_join((nod.combined_filt@meta.data %>% dplyr::select(mouse_id_new) %>% unique), df3)
df4[is.na(df4)] <- 0
df4  <- df4  %>% pivot_longer(!mouse_id_new, values_to = "freq", names_to = "annotations_coarse")

# Control - all sums should be one
# df4 %>% group_by(MouseID_tissue) %>% summarise(sum = sum(freq))

# As we've lost non-grouping variables, let's join them back
md_to_join <- seurat_meta_data %>% 
  dplyr::select(MouseID, Diet, Tissue) %>% 
  unique() %>% 
  mutate(MouseID_tissue = paste(MouseID, Tissue))
df4  <- left_join(df4, md_to_join)

# The final plot
df4 %>% 
  ggplot(aes(x = factor(Diet, labels = c("GFD","STD")), y = freq*100)) + # you can change the x to whatever variable you're interested in
  geom_boxplot(outlier.shape = NA, aes(fill = Diet), alpha = 0.5) +
 geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(shape=21, size = 3, position=position_jitter(width = 0.1, height = 0), 
              aes(fill = Diet)) + theme_classic() + xlab("Group") +  
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5), 
        strip.background = element_blank(),
        axis.text = element_text(color = "black", size = 16),
        axis.ticks.x = element_blank(),
        strip.text = element_text(size = 16)) + 
  scale_color_manual(values = c("dodgerblue","indianred2")) + 
  scale_fill_manual(values = c("dodgerblue","indianred2")) +
  facet_wrap(~annotations_coarse, scales = "free", ncol = 5) +
  ylim(0,NA) +
  ylab("Frequency") +
  xlab("Condition") +
  ggpubr::stat_compare_means(label.x = 1.2, label.y.npc = "top", size = 5, vjust = 0.3, label = "p.format") 

 ggsave("fig/clusters_boxplots_SPL_coarse.png", width = 25, height = 16, units = "cm", limitsize = FALSE)
ggsave("fig/clusters_boxplots_SPL_coarse.svg", width = 25, height = 16, units = "cm", limitsize = FALSE)
```


```{r}
seurat_meta_data <- funda_merge_pln@meta.data


# Create grouped dataframe, calculate the frequencies of clusters
df4 <- seurat_meta_data %>% 
  group_by(MouseID, annotations_coarse) %>% 
  summarise(n = n()) %>% 
  unique() %>% 
  mutate(freq = n / sum(n)) 

df4 <- df4 %>% dplyr::select(-n)  %>% 
ungroup   %>% 
pivot_wider(names_from = "annotations_coarse", values_from = "freq", values_fill = 0) 
df4  <- left_join((seurat_meta_data %>% dplyr::select(MouseID) %>% unique), df4)
df4[is.na(df4)] <- 0
df4  <- df4  %>% pivot_longer(!MouseID, values_to = "freq", names_to = "annotations_coarse")


# Control - all sums should be one
# df4 %>% group_by(MouseID_tissue) %>% summarise(sum = sum(freq))

# As we've lost non-grouping variables, let's join them back
md_to_join <- seurat_meta_data %>% 
  dplyr::select(MouseID, Diet, Tissue) %>% 
  unique() %>% 
  mutate(MouseID_tissue = paste(MouseID, Tissue))
df4  <- left_join(df4, md_to_join)

# The final plot
df4 %>% 
  ggplot(aes(x = factor(Diet, labels = c("GFD","STD")), y = freq*100)) + # you can change the x to whatever variable you're interested in
  geom_boxplot(outlier.shape = NA, aes(fill = Diet), alpha = 0.5) +
 geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(shape=21, size = 3, position=position_jitter(width = 0.1, height = 0), 
              aes(fill = Diet)) + theme_classic() + xlab("Group") +  
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5), 
        strip.background = element_blank(),
        axis.text = element_text(color = "black", size = 16),
        axis.ticks.x = element_blank(),
        strip.text = element_text(size = 16)) + 
  scale_color_manual(values = c("dodgerblue","indianred2")) + 
  scale_fill_manual(values = c("dodgerblue","indianred2")) +
  facet_wrap(~annotations_coarse, scales = "free") +
  ylim(0,NA) +
  ylab("Frequency") +
  xlab("Condition") +
  ggpubr::stat_compare_means(label.x = 1.2, label.y.npc = "top", size = 5, vjust = 0.3, label = "p.format") 

 ggsave("fig/clusters_boxplots_pLN_coarse.png", width = 25, height = 25, units = "cm", limitsize = FALSE)
ggsave("fig/clusters_boxplots_pLN_coarse.svg", width = 25, height = 25, units = "cm", limitsize = FALSE)

```

# Annotations coarse 2 v06

```{r}

md <- funda_merge@meta.data

md <- md %>% mutate(annotations_fine = recode(seurat_clusters, 
                                                     "0" = "CD4 Naive",
                                                     "1" = "CD8 Naive",
                                                     "2" = "CD4 Naive",
                                                     "3" = "CD4 Naive",
                                                     "4" = "CD8 Naive",
                                                     "5" = "CD4 Naive",
                                                     "6" = "CD4 Naive",
                                                     "7" = "CD4 Treg and Tfh",
                                                     "8" = "CD4 Treg and Tfh",
                                                     "9" = "Tgd and NKT ",
                                                     "10" = "NK cells",
                                                     "11" = "CD4 Treg and Tfh",
                                                     "12" = "CD4 Egr Myb",
                                                     "13" = "CD8 ISAGhi",
                                                     "14" = "CD4 Early act. 41BB CD69",
                                                     "15" = "CD8 AIMT",
                                                     "16" = "Proliferating",
                                                     "17" = "ILC",
                                                     "18" = "CD8 Egr"))

funda_merge@meta.data <- md
rownames(funda_merge@meta.data) <- colnames(funda_merge)

```


```{r}
DimPlot(funda_merge, group.by = "annotations_fine", label = T, 
        raster = T, pt.size = 15, raster.dpi = c(5000,5000)) + scale_color_manual(values = c(
  "#6a9bc3", # CD4 Naive
  "#b3e0a6", # CD8 Naive
  
 # "#F28E2B" ,
  "#F28E2B", #Treg
  "#9D7660" , #
  "#B07AA1" ,
  
  "#b9ddf1", # "#F1CE63",
"#24693d",
   "dodgerblue4",
  "#86BCB6" ,
  
  "indianred2",
 "#D4A6C8", 
   "#499894"  ,
 
  "slategray"))



ggsave(filename = "figv06/dimPlot_12clusters.png", width = 20, height = 13, units = "cm")
ggsave(filename = "figv06/dimPlot_12clusters.svg", width = 20, height = 13, units = "cm")

```


# 12 clusters heatmap

## Cluster markers

```{r fig.width = 12, fig.height=4}
markers  <- str_to_sentence(rev(c("TRAC","TRBC1","CD3D","CD4",
                                  "CCR7","SELL",  "TCF7", "LEF1", "IL7R",
                                  "MYC","CD69","TNF","TNFRSF9",
                                  "FOXP3","CTLA4","IL2RA","IL10","CD44",
                                  "CD8A","Isg15","Ifit1","Irf7",
                                  "CXCR3","IL2RB","EOMES","Cd160","Id2",
                                  "TRDC","TRGV2","Cxcr6","Ifng","Xcl1","Kit",
                                  "GZMB","CX3CR1",
                                  "TBX21","KLRG1","KLRK1",
                                  "NCR1","GZMA",
                                  "Fcer1g",
                                  "KLRB1A",
                              "MKI67", "MCM6","CDC6","CDK1"
                  )))


avgexp = AverageExpression(funda_merge, features = markers,
                           return.seurat = F, group.by = "annotations_fine", 
                          assay = "RNA")


options(repr.plot.width = 9.5, repr.plot.height = 2.5)
pheatmap(t(avgexp$RNA)[c(1,6,8,3,2,7,12,9,4,11,5,10),], main = "", 
         scale = "column", cluster_cols = F, cluster_rows = F,
        color=colorRampPalette(c("dodgerblue", "grey95", "indianred2"))(50), 
         border_color = "white",
                  fontsize = 12)
```



```{r}
setwd("G:/48_lab/Project scRNAseq/Scripts Verca/2022-10-17_Funda_NOD/")

options(repr.plot.width = 9.5, repr.plot.height = 2.5)
pheatmap(t(avgexp$RNA)[c(1,6,8,3,2,7,12,9,4,11,5,10),], main = "", 
         scale = "column", cluster_cols = F, cluster_rows = F,
        color=colorRampPalette(c("dodgerblue", "grey95", "indianred2"))(50), 
         border_color = "white",
        filename = "figv06/heatmap_annotations_v06.pdf",
        width = 6, height = 2.2, 
                  fontsize = 9)

```

# Split by mouse ID 

```{r fig.width=20}
funda_merge$MouseID %>% table

DimPlot(funda_merge, split.by = "MouseID", group.by = "Diet", cols = c("dodgerblue","indianred2"))

funda_merge$study <- "Funda"
library(RColorBrewer)

funda_merge@meta.data %>% group_by(MouseID, Tissue, study) %>% tally %>% 
  ggplot(aes(x = Tissue, y = n, fill = MouseID, label = n)) + geom_bar(position="stack", stat="identity") + coord_flip() +
  geom_text(size = 5, position = position_stack(vjust = 0.5)) + theme_classic() + ggtheme() +
  scale_fill_manual(values = c(brewer.pal(n = 5, name = "Blues"),brewer.pal(n = 5, name = "Reds")))

ggsave(filename = "./figv06/mouse_id_counts.svg", width = 32, height = 6, units = "cm")

```

# Boxplots annotations coarse tissue

```{r}

DimPlot(funda_merge,  group.by = "Tissue", cols = c("#5fd3bcff","#4a6ca3ff"), shuffle = T, pt.size = 2)


DimPlot(funda_merge,  group.by = "Tissue", cols = c("#5fd3bcff","#4a6ca3ff"), shuffle = T, 
        raster = T, pt.size = 10, raster.dpi = c(2500,2500)) 
ggsave(filename = "fig/dimPlot_tissue.png", width = 16, height = 13, units = "cm")
ggsave(filename = "fig/dimPlot_tissue.svg", width = 16, height = 13, units = "cm")
```


```{r}
seurat_meta_data <- funda_merge@meta.data

# Create grouped dataframe, calculate the frequencies of clusters
df4 <- seurat_meta_data %>% 
  mutate(MouseID_tissue = paste(MouseID, Tissue)) %>% 
  group_by(MouseID_tissue, annotations_fine) %>% 
  summarise(n = n()) %>% 
  unique() %>% 
  mutate(freq = n / sum(n)) 

# Control - all sums should be one
# df4 %>% group_by(MouseID_tissue) %>% summarise(sum = sum(freq))

# As we've lost non-grouping variables, let's join them back
md_to_join <- seurat_meta_data %>% 
  dplyr::select(MouseID, Diet, Tissue) %>% 
  unique() %>% 
  mutate(MouseID_tissue = paste(MouseID, Tissue))
df4  <- left_join(df4, md_to_join)

# The final plot
df4 %>% 
  ggplot(aes(x = factor(Tissue, labels = c("pLN", "SPL")), y = freq*100)) + # you can change the x to whatever variable you're interested in
  geom_boxplot(outlier.shape = NA, aes(fill = Tissue), alpha = 0.5) +
 geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(shape=21, size = 3, position=position_jitter(width = 0.1, height = 0), 
              aes(fill = Tissue)) + theme_classic() + xlab("Group") +  
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5), 
        strip.background = element_blank(),
        axis.text = element_text(color = "black", size = 16),
        axis.ticks.x = element_blank(),
        strip.text = element_text(size = 16)) + 
  scale_color_manual(values = c("#5fd3bcff","#4a6ca3ff")) + 
  scale_fill_manual(values = c("#5fd3bcff","#4a6ca3ff")) +
  facet_wrap(~annotations_fine, scales = "free", ncol = 6) +
  ylim(0,NA) +
  ylab("Frequency") +
  xlab("Condition") +
  ggpubr::stat_compare_means(label.x = 1.2, label.y.npc = "top", size = 5, vjust = 0.3, label = "p.format") 

 ggsave("figv06/clusters_boxplots_tissue.png", width = 25, height = 12, units = "cm", limitsize = FALSE)
ggsave("figv06/clusters_boxplots_tissue.svg", width = 25, height = 12, units = "cm", limitsize = FALSE) 

```


# Clonal expansion
## Clone abundance

```{r}
metadata_6 <- funda_merge@meta.data
rownames(metadata_6) <- colnames(funda_merge)

metadata_6 <- metadata_6 %>% mutate(
  clone_nt = paste(cdr3_B_nt,cdr3_A1_nt,cdr3_A2_nt),
  clone_aa = paste("CDR3b",cdr3_B,"CDR3a",cdr3_A1)
)

clone_table <- metadata_6 %>%
  dplyr::group_by(factor(clone_aa)) %>%
          dplyr::summarize(n = n(), sum = sum()) %>%
    arrange(desc(n))

clone_table
```


```{r}
metadata_6$test <- 0
metadata_6 <- metadata_6 %>% group_by(test, clone_aa) %>% mutate(clone_abundance = as.numeric(n()))


metadata_6 <- as.data.frame(metadata_6 %>% mutate(clone_abundance = as.numeric(ifelse(clone_abundance==1056,NA_integer_,clone_abundance))) %>%
  mutate(log_clone_abundance = log(clone_abundance, base = 2)))
rownames(metadata_6) <- colnames(funda_merge)

funda_merge@meta.data <- metadata_6
rownames(funda_merge@meta.data) <- colnames(funda_merge)

funda_merge <- AddMetaData(funda_merge, as.numeric(metadata_6$clone_abundance), "clone_abundance")
rownames(funda_merge@meta.data) <- colnames(funda_merge)

FeaturePlot(funda_merge, reduction = "umap", features = "log_clone_abundance")

FeaturePlot(subset(funda_merge, clone_abundance >=1 & clone_abundance < 1056), reduction = "umap", features = "log_clone_abundance", cols = c("lightblue","firebrick")) + theme_classic() + theme(plot.title = element_text(hjust = 0.5)) + ggtitle("Log2 clone abundance")

#ggsave("fig/Log2CloneAbundance.png", width = 13, height = 8, units = "cm", dpi = 120)
#ggsave("fig/Log2CloneAbundance.svg", width = 13, height = 8, units = "cm", dpi = 120)

# Clone abundance bar graph
metadata_6 <- metadata_6 %>%
  mutate( clone_abundance_group = case_when(clone_abundance==5 ~ "5",
                                            clone_abundance==4 ~ "4",
                                            clone_abundance==3 ~ "3",
                                            clone_abundance==2 ~ "2",
                                            clone_abundance==1 ~ "1",
                                            TRUE ~ "0"
                                            ),
          clone_abundance_isNA = if_else(clone_abundance %in% c(1:5), "1", "0")) 

metadata_6 %>% 
  ggplot(aes(x = annotations_fine)) + 
  geom_bar(aes(fill = factor(clone_abundance_group)), position = "fill") + 
  coord_flip() +
  scale_fill_brewer(palette = "Blues") + xlab("Frequency")+
  ylab("") + 
  theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5), 
        legend.position="right", 
        panel.border = element_blank(), 
        legend.title = element_blank()) + 
  ggtitle("Clone abundance in clusters")


```


```{r}
## TCR detection

metadata_6 %>% 
  ggplot(aes(x = factor(annotations_fine, levels =  rev(levels(factor(metadata_6$annotations_fine))[c(1,6,8,3,2,7,12,9,4,11,5,10)])))) + 
  geom_bar(aes(fill = factor(clone_abundance_isNA)), position = "fill") + 
  coord_flip() +
  scale_fill_manual(values = c("grey80", "#c6e9afff")) + xlab("Frequency")+
  ylab("") + 
  theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5), 
        legend.position="right", 
        panel.border = element_blank(), 
        legend.title = element_blank()) + 
  ggtitle("TCR detection in clusters")

ggsave("figv06/TCR_Detection_bar.png", width = 18, height = 10, units = "cm", dpi = 120)
ggsave("figv06/TCR_Detection_bar.svg", width = 18, height = 10, units = "cm", dpi = 120)


```



```{r}

metadata_6$clone_abundance_group
### Clonal abundance
 metadata_6 %>% 
  filter(clone_abundance_isNA == 1) %>% 
  ggplot(aes(x = factor(annotations_fine, levels = rev(levels(factor(metadata_6$annotations_fine))[c(1,6,8,3,2,7,12,9,4,11,5,10)])))) + 
  geom_bar(aes(fill = factor(clone_abundance_group)), position = "fill") + 
   coord_flip() +
  scale_fill_manual(values = c("#eff3ffff", "#bdd7e7ff", "#6baed6ff",
                               "#ce7072ff", "#9c080dff")) + xlab("Frequency")+
  xlab("Frequency") +
  ylab("") + 
  theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5), legend.position="right", panel.border = element_blank(), legend.title = element_blank()) + 
  ggtitle("Clone abundance in clusters") + 
  ggtheme()

 ggsave("figv06/Log2CloneAbundance_bar.png", width = 18, height = 10, units = "cm", dpi = 120)
 ggsave("figv06/Log2CloneAbundance_bar.svg", width = 18, height = 10, units = "cm", dpi = 120)


```

```{r}

DimPlot(funda_merge, cells.highlight = colnames(subset(funda_merge, cdr3_A1 == "CVVGDRGSALGRLHF")))
ggsave(filename = "figv06/dimPlot_iNKT.png", width = 17.5, height = 13, units = "cm")
ggsave(filename = "figv06/dimPlot_iNKT.svg", width = 17.5, height = 13, units = "cm")

```

## Markers table

```{r}
Idents(funda_merge) <- funda_merge$annotations_fine

mrk <- FindAllMarkers(funda_merge, only.pos = T)
```

```{r}
mrk
write.csv(mrk, "240805_markers.csv")
```

```{r}
DimPlot(funda_merge, group.by = "seurat_clusters")

Idents(funda_merge) <- funda_merge$seurat_clusters
```

```{r}
mrk2 <- FindAllMarkers(funda_merge)
write.csv(mrk2, "240615_markers2.csv")

```


## Boxplots diet

```{r fig.width=15, fig.height=5}
seurat_meta_data <- funda_merge@meta.data

# Create grouped dataframe, calculate the frequencies of clusters
df4 <- seurat_meta_data %>% 
  mutate(MouseID_tissue = paste(MouseID, Tissue)) %>% 
  group_by(MouseID_tissue, annotations_coarse) %>% 
  summarise(n = n()) %>% 
  unique() %>% 
  mutate(freq = n / sum(n)) %>% 
  dplyr::select(-n)  %>%
ungroup   %>%
pivot_wider(names_from = "annotations_coarse", values_from = "freq", values_fill = 0) %>% 
  pivot_longer(!MouseID_tissue, names_to = "annotations_coarse", values_to = "freq")

df4
```

### pLN

```{r fig.width=15, fig.height=5}

# Control - all sums should be one
# df4 %>% group_by(MouseID_tissue) %>% summarise(sum = sum(freq))

# As we've lost non-grouping variables, let's join them back
md_to_join <- seurat_meta_data %>% 
  dplyr::select(MouseID, Diet, Tissue) %>% 
  unique() %>% 
  mutate(MouseID_tissue = paste(MouseID, Tissue))
df4  <- left_join(df4, md_to_join)

# The final plot
df4 %>% 
  dplyr::filter(Tissue == "pLN") %>% 
  ggplot(aes(x = Diet, y = freq*100)) + # you can change the x to whatever variable you're interested in
  geom_boxplot(outlier.shape = NA, aes(fill = Diet), alpha = 0.5) + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(position=position_jitter(0.2), size = 2, aes(fill = Diet), shape = 21) + # in aes, you can also use shape or fill (for the shapes that allow it)
  facet_wrap(~factor(annotations_coarse, levels = levels(factor(metadata_6$annotations_coarse))[c(1,6,8,3,2,7,12,9,4,11,5,10)]), scales = "free", ncol = 12) +
  scale_fill_manual(values = c("dodgerblue","indianred2")) + 
  scale_color_manual(values = c("dodgerblue","indianred2")) + 
  ylab("Frequency") +
  xlab("Condition") +
  theme_classic() +
  theme(strip.background = element_blank(), panel.grid = element_blank()) + 
  theme(axis.text.x = element_text(angle = 90)) +
theme(plot.title = element_text(hjust = 0.5), 
        strip.background = element_blank(),
        axis.text = element_text(color = "black", size = 20),
        axis.ticks.x = element_blank(),
        strip.text = element_text(size = 20)) + 
  ylim(0,NA) +
  ylab("") +
  xlab("") +
  ggpubr::stat_compare_means(label.x = 1.2, label.y.npc = "top", size = 5, vjust = 0.3, label = "p.format", method = "t.test") 

 ggsave("figv06/clusters_boxplots_diet_pLN2.png", width = 48, height = 10, units = "cm", limitsize = FALSE)
ggsave("figv06/clusters_boxplots_diet_pLN2.svg", width = 48, height = 10, units = "cm", limitsize = FALSE) 

library(ggpubr)
df5 <- df4 %>% 
  dplyr::filter(Tissue == "pLN") 
compare_means(freq ~ Diet, data = df5, 
              group.by = "annotations_coarse",
              method = "t.test")


compare_means(freq ~ Diet, data = df5, 
              group.by = "annotations_coarse")

write.csv(df5, "pLN_freq.csv")
```

### Spleen

```{r fig.width=15, fig.height=5}
# Control - all sums should be one
# df4 %>% group_by(MouseID_tissue) %>% summarise(sum = sum(freq))

# The final plot
df4 %>% 
  dplyr::filter(Tissue == "Spleen") %>% 
  ggplot(aes(x = Diet, y = freq*100)) + # you can change the x to whatever variable you're interested in
  geom_boxplot(outlier.shape = NA, aes(fill = Diet), alpha = 0.5) + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(position=position_jitter(0.2), size = 2, aes(fill = Diet), shape = 21) + # in aes, you can also use shape or fill (for the shapes that allow it)
  facet_wrap(~factor(annotations_coarse, levels = levels(factor(metadata_6$annotations_coarse))[c(1,6,8,3,2,7,12,9,4,11,5,10)]), scales = "free", ncol = 12) +
  scale_fill_manual(values = c("dodgerblue","indianred2")) + 
  scale_color_manual(values = c("dodgerblue","indianred2")) + 
  ylab("Frequency") +
  xlab("Condition") +
  theme_classic() +
  theme(strip.background = element_blank(), panel.grid = element_blank()) + 
  theme(axis.text.x = element_text(angle = 90)) +
theme(plot.title = element_text(hjust = 0.5), 
        strip.background = element_blank(),
        axis.text = element_text(color = "black", size = 20),
        axis.ticks.x = element_blank(),
        strip.text = element_text(size = 20)) + 
  ylim(0,NA) +
  ylab("") +
  xlab("") +
  ggpubr::stat_compare_means(label.x = 1.2, label.y.npc = "top", size = 5, vjust = 0.3, label = "p.format", method = "t.test") 

ggsave("figv06/clusters_boxplots_diet_SPL2.png", width = 48, height = 10, units = "cm", limitsize = FALSE)
ggsave("figv06/clusters_boxplots_diet_SPL2.svg", width = 48, height = 10, units = "cm", limitsize = FALSE) 


library(ggpubr)
df5 <- df4 %>% 
  dplyr::filter(Tissue == "Spleen") 
compare_means(freq ~ Diet, data = df5, 
              group.by = "annotations_coarse",
              method = "t.test")


compare_means(freq ~ Diet, data = df5, 
              group.by = "annotations_coarse")

write.csv(df5, "SPL_freq.csv")

```

# DE GFD vs GTD

```{r}
DimPlot(funda_merge, group.by = "Diet", cols = c("#92c0df", "indianred3"), shuffle = T)


metadata_6 %>% 
  ggplot(aes(x = factor(annotations_fine, levels =  rev(levels(factor(metadata_6$annotations_fine))[c(1,6,8,3,2,7,12,9,4,11,5,10)])))) + 
  geom_bar(aes(fill = factor(Diet)), position = "fill") + 
  coord_flip() +
  scale_fill_manual(values = c("dodgerblue", "indianred2")) + xlab("Frequency")+
  ylab("") + 
  theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5), 
        legend.position="right", 
        panel.border = element_blank(), 
        legend.title = element_blank()) + 
  ggtitle("TCR detection in clusters")



```




```{r}
Idents(funda_merge) <- funda_merge$Diet


mrk_de_whole_data <- FindAllMarkers(funda_merge, only.pos = TRUE, logfc.threshold = 0.1, min.pct = 0.05)
mrk_de_whole_data


mrk_de_whole_data %>% filter(cluster == "Standard") %>% nrow
```

Spleen vs pLN

```{r}
funda_merge$Tissue %>% table
mrk_de_pLN <- FindAllMarkers(subset(funda_merge, Tissue == "pLN"), only.pos = TRUE, logfc.threshold = 0.1, min.pct = 0.05)
mrk_de_spl <- FindAllMarkers(subset(funda_merge, Tissue == "Spleen"), only.pos = TRUE, logfc.threshold = 0.1, min.pct = 0.05)

mrk_de_pLN
mrk_de_spl

dir.create("tables")
write.csv(mrk_de_whole_data, "tables/mrk_de_whole_data.csv")
write.csv(mrk_de_pLN, "tables/mrk_de_pLN.csv")
write.csv(mrk_de_spl, "tables/mrk_de_spl.csv")

```


```{r}
library(ggVennDiagram)
library(gplots)
```


```{r}
x_gfd <- list(
  pLN = mrk_de_pLN %>% dplyr::filter(cluster == "GFD") %>% pull(gene), 
  SPL = mrk_de_spl %>% dplyr::filter(cluster == "GFD") %>% pull(gene)) 
 ggVennDiagram(x_gfd)
  ggsave("figs/venn_gfd.svg", width = 10, height = 6, units = "cm")

 
 v.table <- venn(x_gfd)
 print(v.table)
```

```{r}
x_std <- list(
  pLN = mrk_de_pLN %>% dplyr::filter(cluster == "Standard") %>% pull(gene), 
  Spl = mrk_de_spl %>% dplyr::filter(cluster == "Standard") %>% pull(gene)) 
 ggVennDiagram(x_std)
 ggsave("figs/venn_std.svg", width = 10, height = 6, units = "cm")
 v.table <- venn(x_std)

    print(v.table)
```


```{r}
library(ensembldb)
library(EnsDb.Mmusculus.v79)

geneIDs1 <- ensembldb::select(EnsDb.Mmusculus.v79, keys= mrk_de_whole_data$gene, 
                              keytype = "SYMBOL", columns = c("SYMBOL","ENTREZID"))

colnames(geneIDs1)  <- c("gene", "entrezid")

geneIDs1

all_markers2  <- mrk_de_whole_data  %>% 
                     left_join(geneIDs1) 

all_markers2
```

## Heatmap all DE genes


```{r}
unique(mrk_de_whole_data$gene)

genes_selection <- c("S100a6", "Lef1","Irf7","Lck","Lat","Id3", "Myc", "Itgb2", "Il21r", "Ptpn6","Cd40lg",        "Ifngr1",
"Foxp3","Ly6a", "Usp18","Klrd1","Ifi27","Isg15" ,"Foxo1","Bcl2","Epsti1","Slamf6","Il18r1")


rownames(avgexp$RNA)
```

```{r}
genes_selection <- c("Lef1","Irf7","Lck","Lat","Id3", "Myc",  "Cd40lg", "Ifngr1",
"Foxp3","Ly6a", "Usp18","Klrd1","Ifi27","Isg15" ,"Foxo1","Epsti1","Slamf6","Il18r1","Ddit4","Sgk1")

```


```{r}
library(ComplexHeatmap)
library(circlize)

ha = rowAnnotation(foo = anno_mark(at = which(rownames(avgexp$RNA) %in% genes_selection), 
    labels = rownames(avgexp$RNA)[which(rownames(avgexp$RNA) %in% genes_selection)]))


scaled_m <- scale(t(avgexp$RNA))

library(colorspace)
Heatmap(t(scaled_m), col = colorRamp2(c(-2, 0, 2), c("dodgerblue", "white", "indianred2")), 
    name = "scaled_expr", 
    show_column_names = TRUE, width = unit(8, "cm"), right_annotation = ha, show_row_names = FALSE) 

pdf("./figs/heatmap_gfd_vs_std.pdf", width = 3, height = 5)
Heatmap(t(scaled_m), col = colorRamp2(c(-2, 0, 2), c("dodgerblue", "white", "indianred2")), 
    name = "scaled_expr", 
    show_column_names = TRUE, width = unit(8, "cm"), right_annotation = ha, show_row_names = FALSE) 
dev.off()

```


## GSEA

```{r}
gsea_upGFD  <- all_markers2  %>% 
                        dplyr::filter(!is.na(entrezid) & cluster == "GFD")  %>% 
                        group_by(entrezid, avg_log2FC)  %>% 
                        tally()  %>% 
                        arrange(desc(n))  %>% 
                        pull(entrezid)
gsea_upGFD

gsea_upSTD  <- all_markers2  %>% 
                        dplyr::filter(!is.na(entrezid) & cluster == "Standard")  %>% 
                        group_by(entrezid, avg_log2FC)  %>% 
                        tally()  %>% 
                        arrange(desc(n))  %>% 
                        pull(entrezid)

gsea_upSTD
```


```{r}
library(msigdbr)

immune_genes <- immune_genes <- msigdbr(species = "Mus musculus", category = "C7") %>% 
  dplyr::select(gs_name, entrez_gene)
head(immune_genes)


which(all_markers2$entrezid %in% immune_genes)

all_markers2$gene[which(all_markers2$entrezid %in% immune_genes$entrez_gene)]

```



```{r}
geneList_nod <- all_markers2 %>% filter(!is.na(entrezid) & cluster == "GFD")
write.csv(geneList_nod, "geneList_nod.csv")


geneList_nod <- all_markers2 %>% filter(!is.na(entrezid) & cluster == "Standard")
write.csv(geneList_nod, "geneList_nod_std.csv")


geneList_nod_vector <- geneList_nod$avg_log2FC
names(geneList_nod_vector) <- geneList_nod$entrezid

geneList_nod_vector <- geneList_nod_vector[rev(order(geneList_nod_vector))]
geneList_nod_vector


edo2 <- gseDO(geneList_nod_vector)

p1 <- gseaplot(edo2, geneSetID = 1, by = "runningScore", title = edo2$Description[1])


em2 <- GSEA(geneList_nod_vector, TERM2GENE = immune_genes)
em %>% as.data.frame()


mutate(em, qscore = -log(p.adjust, base=10)) %>% 
    barplot(x="qscore")

dotplot(em, showCategory=20) + ggtitle("dotplot for ORA")
```

```{r}
em <- enricher(gsea_upSTD, TERM2GENE=immune_genes)
head(em)
em %>% as.data.frame()

gsea_upSTD

dotplot(em, showCategory=20) + ggtitle("dotplot for ORA")
```


```{r}

all_markers2
gsea_upSTD  <- all_markers2  %>% 
                        dplyr::filter(!is.na(entrezid) & cluster == "Standard")  %>% 
                        group_by(entrezid, avg_log2FC)  %>% 
                        tally()  %>% 
                        arrange(desc(n))  %>% 
                        pull(entrezid)
gsea_upSTD


ego2 <- enrichGO(gene          = gsea_upSTD,
                OrgDb         = org.Mm.eg.db,
                ont           = "ALL",
                pAdjustMethod = "BH",
                pvalueCutoff  = 0.01,
                qvalueCutoff  = 0.05,
        readable      = TRUE)


head(ego2)

ego2 %>% as.data.frame
```


## GSEA traditional

Get fold changes using marker testing with unlimited thresholds.

```{r}
Idents(funda_merge) %>% table

fc.diet <- FindAllMarkers(funda_merge, test.use = "wilcox", logfc.threshold = -Inf, only.pos = F, pseudocount.use = 10^-3, min.cells.feature = 1, min.pct = -Inf, return.thresh = 1)
```

Prepare a list with immune pathways.

```{r}
c7.pathways <- msigdbr(species = "Mus musculus", category = "C7")
c7.pathways.names <- levels(factor(c7.pathways$gs_name))


fct_generate_pathway <- function(i){
  pathway <- c7.pathways %>% dplyr::filter(gs_name == c7.pathways.names[i])
  pathway_vector <- c(c7.pathways.names[i], pull(pathway, "gene_symbol"))
}

pathways <- map(.x = 1:length(c7.pathways.names), .f = fct_generate_pathway)

pathways

pathways2 <- list()

for(i in 1:length(c7.pathways.names)){
  pathways2[[i]] <- pathways[[i]][2:length(pathways[[i]])]
  names(pathways2)[i] <- pathways[[i]][1]
}

```


Get fold changes for GFD vs STD. Calculate fGSEA for all pathways.

```{r fig.height = 13, fig.width=20}

fc.diet.gfd <- fc.diet %>% filter(cluster == "GFD")
fc.diet.gfd <- data.frame(avg_logFC = fc.diet.gfd$avg_log2FC, gene = fc.diet.gfd$gene)
fc.genes <- as.vector(fc.diet.gfd$gene)                                

fc.entrezids <- AnnotationDbi::select(org.Mm.eg.db, 
       keys = fc.genes,
       columns = c("ENTREZID", "SYMBOL"),
       keytype = "SYMBOL")

fc.diet.gfd$entrezid <- fc.entrezids$ENTREZID

fc.diet.gfd


fc.input <- as.numeric(fc.diet.gfd$avg_logFC)
names(fc.input) <- fc.diet.gfd$gene


fgseaRes.gfd <- fgsea(pathways = pathways2, 
                  stats    = fc.input)

fgseaRes.gfd %>% arrange(pval)

fgseaRes.gfd.signif <- fgseaRes.gfd %>% filter(padj < 0.05)


fgseaRes.gfd.signif %>% arrange(pval)


fgseaRes.gfd.signif %>% 
  mutate(pathway = fct_reorder(pathway, NES, .fun='median')) %>%
  ggplot(aes(x = NES, y = pathway)) +
  geom_point(aes(size = size, color = padj)) + 
  scale_color_gradient(low = "red2", high = "lightblue") + 
  theme_classic() +
  theme(axis.ticks.y = element_blank()) + ggtheme()
ggsave(filename = "figs/pathway_plot_big.png", width = 35, height = 28, units = "cm")
ggsave(filename = "figs/pathway_plot_big.svg", width = 35, height = 28, units = "cm")


ggsave(filename = "figs/pathway_plot_small.png", width = 35, height = 10, units = "cm")



```


```{r}
#  GSE14415_INDUCED_TREG_VS_TCONV_UP

fgsea_2 <- fgsea(pathways = list(our_list = pathways2$GSE14415_INDUCED_TREG_VS_TCONV_UP), 
                                      stats    = fc.input)
p1 <- plotEnrichment(pathways2$GSE14415_INDUCED_TREG_VS_TCONV_UP, fc.input) + ylim(-0.1, 0.55) +  labs(title="\nGSE14415_INDUCED_TREG_VS_TCONV_UP", 
    subtitle = paste0("p-value = ",format.pval(fgsea_2$pval, digits = 2)))
p1


#  GSE28726_NAIVE_VS_ACTIVATED_CD4_TCELL_DN

fgsea_2 <- fgsea(pathways = list(our_list = pathways2$GSE28726_NAIVE_VS_ACTIVATED_CD4_TCELL_DN), 
                                      stats    = fc.input)
p2 <- plotEnrichment(pathways2$GSE28726_NAIVE_VS_ACTIVATED_CD4_TCELL_DN, fc.input) + ylim(-0.1, 0.55) + labs(title="\n GSE28726_NAIVE_VS_ACTIVATED_CD4_TCELL_DN", 
    subtitle = paste0("p-value = ",format.pval(fgsea_2$pval, digits = 2)))
p2

#  GOLDRATH_NAIVE_VS_EFF_CD8_TCELL_DN

fgsea_2 <- fgsea(pathways = list(our_list = pathways2$GOLDRATH_NAIVE_VS_EFF_CD8_TCELL_DN), 
                                      stats    = fc.input)
p3 <- plotEnrichment(pathways2$GOLDRATH_NAIVE_VS_EFF_CD8_TCELL_DN, fc.input) + ylim(-0.1, 0.55) + labs(title="\n GOLDRATH_NAIVE_VS_EFF_CD8_TCELL_DN", 
    subtitle = paste0("p-value = ",format.pval(fgsea_2$pval, digits = 2)))
p3

cowplot::plot_grid(p1, p2, p3, ncol = 1)
ggsave(filename = "figs/gsea.svg", width = 10, height = 20, units = "cm")

```



# Correaltion with FACS

```{r}
DimPlot(funda_merge, group.by = "annotations_coarse")


funda_merge@meta.data <- funda_merge@meta.data %>% mutate(
  annotations_facs = case_when(annotations_coarse %in% c("CD4 Naive","CD4 Early activated") ~ "CD4_naive",
                               annotations_coarse %in% c("CD4 Treg", "Proliferating" ) ~ "Treg_and_CD4_eff",
                                annotations_coarse %in% c("CD8 Early activated", "CD8 Naive") ~ "CD8_naive",
                               annotations_coarse %in% c("Tgd and NKT", "ILC") ~ "Tgd_NKT",
                               annotations_coarse %in% c("CD8 AIMT") ~ "CD8_AIMT_and_eff",
                               annotations_coarse %in% c("NK cells") ~ "NK_cells"
                               )) %>% 
  mutate(annotations_facs = ifelse(seurat_clusters == 4, "CD8_AIMT_and_eff",annotations_facs))

DimPlot(funda_merge, group.by = "annotations_facs")

seurat_meta_data <- funda_merge@meta.data
seurat_meta_data$Tissue_ID <- paste(seurat_meta_data$Tissue, seurat_meta_data$MouseID, sep = ".")

# Create grouped dataframe, calculate the frequencies of clusters
df4 <- seurat_meta_data %>% 
  group_by(Tissue_ID, annotations_facs) %>% 
  summarise(n = n()) %>% 
  unique() %>% 
  mutate(freq = n / sum(n)) 

# Control - all sums should be one
# df4 %>% group_by(MouseID) %>% summarise(sum = sum(freq))

# As we've lost non-grouping variables, let's join them back
md_to_join <- seurat_meta_data %>% 
  dplyr::select(Tissue_ID, MouseID, Diet, Tissue) %>% 
  unique()

df4  <- left_join(df4, md_to_join)

df4
```


```{r}
df4
df5  <- df4 %>% 
ungroup   %>% 
  mutate(Tissue_ID_Diet = paste(Tissue_ID, Diet, sep = ".")) %>% 
  dplyr::select(Tissue_ID_Diet, freq, annotations_facs) %>% 
pivot_wider(names_from = "Tissue_ID_Diet", values_from = "freq", values_fill = 0) 


df6  <- df5 %>% pivot_longer(!annotations_facs, names_to = "Tissue_ID_Diet", values_to = "freq") %>% 
  separate(Tissue_ID_Diet, into = c("Tissue","MouseID","Diet"), sep = "\\.", remove = F)

df6
```


```{r fig.height = 5.5, fig.width = 8}

df6 %>% 
  dplyr::filter(Tissue == "Spleen") %>% 
  ggplot(aes(x = Diet, y = freq*100)) + # you can change the x to whatever variable you're interested in
  geom_boxplot(outlier.shape = NA, aes(fill = Diet), alpha = 0.7) + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(width = 0.1, height = 0.1, shape = 21,
              size = 4, aes(fill = Diet)) + # in aes, you can also use shape or fill (for the shapes that allow it)
  facet_wrap(~annotations_facs, scales = "free", ncol = 6) +
  scale_fill_manual(values = c("dodgerblue","indianred2")) + 
  scale_color_manual(values = c("dodgerblue","indianred2")) + 
  ylab("Frequency") +
  xlab("Condition") +
  theme_classic() +
  ylim(0,NA)+
  theme(strip.background = element_blank(), panel.grid = element_blank()) + 
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) + ggtitle("Spleen scRNAseq") +
  ggpubr::stat_compare_means(label.x = 1.2, label.y.npc = "top", size = 5, vjust = 0.3, label = "p.format") + 
  ggtheme()

ggsave("figs/spl_scRNAseq.svg", width = 36, height = 9.5, units = "cm")

df6 %>% 
  dplyr::filter(Tissue == "pLN") %>% 
  ggplot(aes(x = Diet, y = freq*100)) + # you can change the x to whatever variable you're interested in
  geom_boxplot(outlier.shape = NA, aes(fill = Diet), alpha = 0.7) + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(width = 0.1, height = 0.1, shape = 21,
              size = 4, aes(fill = Diet)) + # in aes, you can also use shape or fill (for the shapes that allow it)
  facet_wrap(~annotations_facs, scales = "free", ncol = 6) +
  scale_fill_manual(values = c("dodgerblue","indianred2")) + 
  scale_color_manual(values = c("dodgerblue","indianred2")) + 
  ylab("Frequency") +
  xlab("Condition") +
  theme_classic() +
  ylim(0,NA)+
  theme(strip.background = element_blank(), panel.grid = element_blank()) + 
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) + ggtitle("pLN scRNAseq") +
  ggpubr::stat_compare_means(label.x = 1.2, label.y.npc = "top", size = 5, vjust = 0.3, label = "p.format") + 
  ggtheme()


ggsave("figs/pln_scRNAseq.svg", width = 36, height = 9.5, units = "cm")


DimPlot(funda_merge, group.by = "annotations_facs", cols = c("#006bd6ff","#41a90eff","#6ff86fff",
                                                             "#be66beff","#a83e09ff","#ffa448ff"),
        pt.size = 1) 
```

```{r}
df4
```

Load facs percentages

```{r}
high_dim <- read_excel("G:/48_lab/Project Funda NOD/FACS/220614 NOD gluten free diet scRNAseq/HighDim/240125_results.xlsx")


high_dim <- t(high_dim) %>% as.data.frame() %>% rownames_to_column("sample")

high_dim <- high_dim[2:301,]
high_dim
colnames(high_dim) <- c("population", "value")

high_dim
high_dim$population <- sub(pattern = "X-shift_Angu_K102_xw0krs_", 
                           replacement = "cluster",
                            x = high_dim$population)
high_dim$population <- sub(pattern = "X- | Freq. of Parent", 
                           replacement = "",
                            x = high_dim$population)


high_dim <- high_dim %>% separate(population, into = c("tissue","sample","cluster"), 
                                  sep = "/", remove = F) %>% mutate(
                                    value = sub(value, pattern = ",", replacement = ".")
                                  ) %>% mutate(value = as.numeric(value))

high_dim <- high_dim %>% filter(!is.na(cluster))


high_dim

high_dim2 <- high_dim %>% dplyr::select(-population) %>% 
  pivot_wider(names_from = "cluster", values_from = "value", values_fill = 0)

high_dim2$Diet <- if_else(high_dim2$sample %in% c(1:5), "GFD", "STD")
high_dim2

high_dim2$tissue <- case_when(high_dim2$tissue == 1 ~ "pLN",
                              high_dim2$tissue == 2 ~ "iLN",
                              high_dim2$tissue == 3 ~ "SPL"
                              )



high_dim2 <- high_dim2 %>% mutate(
  CD8_naive = `cluster1 |`+`cluster8 |`,
  CD8_AIMT_and_eff = `cluster3 |`,
  CD4_naive = `cluster5 |`,
  NK_cells = `cluster9 |`,
  Treg_and_CD4_eff = `cluster6 |`+`cluster7 |`,
  Tgd_NKT = `cluster2 |`+`cluster4 |`+`cluster10 |`
)

high_dim2
```

```{r fig.height = 3, fig.width = 8}


high_dim2 %>% dplyr::select(1,2,13,14:19) %>% 
  pivot_longer(cols = c(4:9), names_to = "annot", values_to = "freq") %>% 
  dplyr::filter(tissue == "pLN") %>% 
  ggplot(aes(x = Diet, y = freq)) + # you can change the x to whatever variable you're interested in
  geom_boxplot(outlier.shape = NA, aes(fill = Diet), alpha = 0.7) + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(width = 0.1, height = 0.1, size = 3, shape = 21, 
              aes(fill = Diet)) + # in aes, you can also use shape or fill (for the shapes that allow it)
  facet_wrap(~annot, scales = "free", ncol = 6) +
  scale_color_manual(values = c("dodgerblue","indianred2")) + 
  scale_fill_manual(values = c("dodgerblue","indianred2")) + 
  ylab("Frequency") +
  xlab("Condition") +
  ylim(0,NA)+
  ggpubr::stat_compare_means(label.x = 1.5, label.y.npc = "top", size = 5, vjust = 0.3, label = "p.format") + 
    theme_classic() +
  theme(strip.background = element_blank(), panel.grid = element_blank()) + 
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())  + ggtitle("pLN FACS") + ggtheme()
  
ggsave("figs/pln_facs.svg", width = 36, height = 9.5, units = "cm")


high_dim2 %>% dplyr::select(1,2,13,14:19) %>% 
  pivot_longer(cols = c(4:9), names_to = "annot", values_to = "freq") %>% 
  dplyr::filter(tissue == "SPL") %>% 
  ggplot(aes(x = Diet, y = freq)) + # you can change the x to whatever variable you're interested in
  geom_boxplot(outlier.shape = NA, aes(fill = Diet), alpha = 0.7) + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(width = 0.1, height = 0.1, size = 3, shape = 21, 
              aes(fill = Diet)) + # in aes, you can also use shape or fill (for the shapes that allow it)
  facet_wrap(~annot, scales = "free", ncol = 6) +
  scale_color_manual(values = c("dodgerblue","indianred2")) + 
  scale_fill_manual(values = c("dodgerblue","indianred2")) + 
  ylab("Frequency") +
  xlab("Condition") +
  ylim(0,NA)+
  theme_classic() +
  ggpubr::stat_compare_means(label.x = 1.5, label.y.npc = "top", size = 5, vjust = 0.3, label = "p.format") + 
    theme(strip.background = element_blank(), panel.grid = element_blank()) + 
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())  + ggtitle("Spleen FACS") + ggtheme()

ggsave("figs/spl_facs.svg", width = 36, height = 9.5, units = "cm")

```

```{r}
high_dim2
dir.create("./results_240618_tissue/")
for(i in c(14:19)){
  
  df_filt <- high_dim2 %>% 
    dplyr::select(1,2,13,i)
  colnames(df_filt) <- c(colnames(df_filt[c(1:3)]), "value")
  df_filt$value <- as.numeric(df_filt$value)

plot <- df_filt %>% 
  ggplot(aes(x = tissue,
             y = value)) +
  geom_boxplot(outlier.shape = NA, aes(fill = tissue), alpha = 0.5) +
 geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(shape=21, size = 3, position=position_jitter(width = 0.1, height = 0), 
              aes(fill = tissue)) + theme_classic() + xlab("Group") +  
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5), 
        strip.background = element_blank(),
        axis.text = element_text(color = "black", size = 16),
        strip.text = element_text(size = 16),
        axis.ticks.x = element_blank()) + 
  scale_color_manual(values = c("#56ac00ff","#5fd3bcff","#4a6ca3ff")) + 
  scale_fill_manual(values = c("#56ac00ff","#5fd3bcff","#4a6ca3ff")) +
  ylim(0,NA) +
  xlab("") + ylab("Value") +
  ylim(0,NA)+
  ggpubr::stat_compare_means(label.x = 1.5, label.y.npc = "top", size = 5, vjust = 0.3, label = "p.format") + 
      theme(plot.title = element_text(hjust = 0.5, size = 22),
          axis.line = element_line(colour = "black"), 
        axis.ticks = element_line(colour = "black")) +
  plot_annotation(
    title = paste(stringr::str_replace_all(string = colnames(high_dim2)[i], pattern = "/", replacement = "\n")) ,
    theme = theme(plot.title = element_text(size = 8), plot.subtitle = element_text(size = 8)))
print(plot)
  ggsave(paste0("./results_240618_tissue/",paste0("plot_",i),".svg"), width = 10, height = 9, units = "cm")
}

getwd()
```

## Correlation

```{r}
df4

high_dim3 <- high_dim2 %>% mutate(Tissue_ID = paste0(tissue, ".Mouse_",sample)) %>% dplyr::select(14:20)
high_dim3
corr_df <- df4 %>% 
  dplyr::select(1,freq, annotations_facs) %>% 
  mutate(freq = freq*100) %>% 
  pivot_wider(names_from = "annotations_facs", values_from = freq, 
              values_fill = 0) 

colnames(corr_df)[2:7] <- paste("FACS",colnames(corr_df)[2:7])

high_dim3$Tissue_ID <- gsub(high_dim3$Tissue_ID, pattern = "SPL", replacement = "Spleen")
colnames(high_dim3)[1:6] <- paste("scRNAseq",colnames(high_dim3)[1:6])
corr_df <-  corr_df %>% left_join(high_dim3)
corr_df

corr_df2 <- corr_df %>% pivot_longer(!Tissue_ID, names_to = "population", values_to = "freq") %>% 
  separate(Tissue_ID, into = c("Tissue","MouseID"), sep = "\\.", remove = F) %>% 
  separate(population, into = c("Method","Population"), sep = " ", remove = T) %>% 
  pivot_wider(names_from = Method, values_from = freq)
  
corr_df2


library(ggpubr)
options(repr.plot.width = 5, repr.plot.height = 5)
  corr_df2 %>%  ggplot(aes(x=FACS, y=scRNAseq)) +
  geom_point(size = 1.5, aes(shape = Tissue, color = Population)) +
 geom_smooth(method=lm) + ggtitle("") +stat_cor() + theme_classic() + ggtheme() +
 scale_y_continuous() + 
    scale_color_manual(values = c("#006bd6ff","#41a90eff","#6ff86fff",
                                                             "#be66beff","#a83e09ff","#ffa448ff")) + 
    ggtheme() + NoLegend()
    ggsave("figs/correlation.svg", width = 7, height = 7, units = "cm")

 
  
  options(repr.plot.width = 5, repr.plot.height = 5)
  corr_df2 %>%  ggplot(aes(x=FACS, y=scRNAseq)) +
  geom_point(size = 1.5, aes(shape = Tissue, color = Population)) +
 geom_smooth(method=lm) + ggtitle("") +stat_cor() + theme_classic() + ggtheme() +
 scale_y_continuous() + facet_wrap(~Population, ncol = 3, scales = "free") + 
    scale_color_manual(values = c("#006bd6ff","#41a90eff","#6ff86fff",
                                                             "#be66beff","#a83e09ff","#ffa448ff"))
  
  
 DimPlot(funda_merge, group.by = "annotations_facs", cols = c("#006bd6ff","#41a90eff","#6ff86fff",
                                                             "#be66beff","#a83e09ff","#ffa448ff")) 

  
```


```{r}
high_dim2

for(i in c(14:19)){
  
  df_filt <- high_dim2 %>% 
    dplyr::select(1,2,13,i)
  colnames(df_filt) <- c(colnames(df_filt[c(1:3)]), "value")
  df_filt$value <- as.numeric(df_filt$value)

plot <- df_filt %>% 
  ggplot(aes(x = Diet,
             y = value)) +
  facet_wrap(~tissue, ncol = 3, scales = "free") +
   geom_dotplot(binaxis='y', stackdir='center', dotsize = 0) + 
   geom_boxplot(outlier.shape = NA) +
    geom_jitter(binaxis='y', position=position_jitter(width = 0.1, height = 0.01), 
                size = 3, stackdir='center', aes(color = Diet)) + 
  theme_classic() + xlab("") +  
 xlab("") + ylab("Value") +
   scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  ggpubr::stat_compare_means(label.x = 1.2, label.y.npc = "top", size = 5, vjust = 0.3, label = "p.format") + 
      theme(plot.title = element_text(hjust = 0.5, size = 22),
          axis.line = element_line(colour = "black"), 
        axis.ticks = element_line(colour = "black")) +
  ylim(0,NA)+
  plot_annotation(
    title = paste(stringr::str_replace_all(string = colnames(high_dim2)[i], pattern = "/", replacement = "\n")) ,
    theme = theme(plot.title = element_text(size = 8), plot.subtitle = element_text(size = 8)))
print(plot)
  #ggsave(paste0("./results_240129/",paste0("plot_",i),".png"), width = 32, height = 12, units = "cm")
}

getwd()
```

# FACS heatmap

## Cluster heatmap

```{r}
hm <- read_xlsx("../../../Project Funda NOD/FACS/Final_analysis/HighDim/240618_cluster_heatmap_6clusters.xlsx")

hm_map <- hm %>% column_to_rownames("Cluster") %>% as.matrix()

hm_map

library(pheatmap)

pheatmap(hm_map, main = "", 
         scale = "column", cluster_cols = T, cluster_rows = T,
        color=colorRampPalette(c("dodgerblue", "grey95", "indianred2"))(50), 
         border_color = "white",
                  fontsize = 12)

pheatmap(hm_map[c(1,4,2,5,3,6),c(10,11,3,5,4,7,8,1,2,9)], main = "", 
         scale = "column", cluster_cols = F, cluster_rows = T,
        color=colorRampPalette(c("dodgerblue", "grey95", "indianred2"))(50), 
         border_color = "white",
                  fontsize = 12, width = 4, height = 3)


pheatmap::pheatmap(hm_map[c(1,4,2,5,3,6),c(10,11,3,5,4,7,8,1,2,9)], main = "", 
         scale = "column", cluster_cols = F, cluster_rows = T,
        color=colorRampPalette(c("dodgerblue", "grey95", "indianred2"))(50), 
         border_color = "white",
                  fontsize = 12, width = 4, height = 3,
        filename = "heatmap_cluster_FACS_3.pdf")

getwd()
```


```{r}
hm <- read_xlsx("../../../220614 NOD gluten free diet scRNAseq/HighDim/cluster_heatmap_coarse.xlsx")

hm_map <- hm %>% column_to_rownames("Cluster") %>% as.matrix()

hm_map

library(pheatmap)

pheatmap(hm_map[,c(10,3,5,6,11,1,2,7,8,9,4)], main = "", 
         scale = "column", cluster_cols = F, cluster_rows = F,
        color=colorRampPalette(c("dodgerblue", "grey95", "indianred2"))(50), 
         border_color = "white",
                  fontsize = 12)


pheatmap(hm_map[c(1,5,3,4,2),c(10,3,5,6,11,1,2,7,8,9,4)], main = "", 
         scale = "column", cluster_cols = F, cluster_rows = F,
        color=colorRampPalette(c("dodgerblue", "grey95", "indianred2"))(50), 
         border_color = "white",
                  fontsize = 12, width = 4, height = 3,
        filename = "heatmap_cluster_FACS.pdf")
```





# IREA

```{r}
## Read IREA table outputs
irea <- list.files("IREA", full.names = T)


for(i in 1:10){
  ire_sample <- read_csv(irea[i])
  ire_sample$i <- irea[i]
  if(i == 1){
      irea_all <- ire_sample
    } else {
      irea_all <- rbind(irea_all, ire_sample)
      }
}

irea_all$...1 <- NULL

irea_all$i <- sub(irea_all$i, pattern = "IREA/df_irea_", replacement = "")
irea_all$i <- sub(irea_all$i, pattern = ".csv", replacement = "")
  
irea_all <-  irea_all %>% separate(i, into = c("Diet", "Cluster"), sep = "_")

irea_all
```


```{r}
irea_all %>% 
   filter(Diet == "gf" & Cluster %in% c("cd4","cd8","treg")) %>% 
  filter(padj < 0.05) %>% 
 ggplot(aes(x = fct_reorder(Cytokine, `Enrichment Score`, .fun='median', .desc = TRUE), y = `Enrichment Score`, fill=..y..)) + 
 stat_summary(fun=mean, geom="bar", position = position_dodge(0.9)) +
      theme_bw()  + scale_size_continuous(range=c(1,4), limits = c(0,20)) +
  geom_point(aes(size = -log(padj, base = 10), shape = Cluster)) + 
  scale_shape_manual(values = c(17,19,15))+
 scale_fill_gradient2(
    low = "dodgerblue3",
  mid = "grey88",
  high = "indianred2",
  limits = c(-750,2250))+xlab("")+
 theme(axis.text.x = element_text(angle = 90, color = "black"),
       axis.ticks.x = element_blank()
       ) + ggtheme()

#ggsave(filename = "figs/irea_gfd.svg", width = 19, height = 10, units = "cm")
#ggsave(filename = "figs/irea_gfd.png", width = 19, height = 10, units = "cm")

```


No filtration pval


```{r}
irea_all %>% 
   filter(Diet == "gf" & Cluster %in% c("cd4","cd8","treg")) %>% 
 # filter(padj < 0.05) %>% 
 ggplot(aes(x = fct_reorder(Cytokine, `Enrichment Score`, .fun='median', .desc = TRUE), y = `Enrichment Score`, fill=..y..)) + 
 stat_summary(fun=mean, geom="bar", position = position_dodge(0.9)) +
      theme_bw()  + scale_size_continuous(range=c(1,4), limits = c(0,20)) +
  geom_point(aes(size = -log(padj, base = 10), shape = Cluster)) + 
  scale_shape_manual(values = c(17,19,15))+
 scale_fill_gradient2(
    low = "dodgerblue3",
  mid = "grey88",
  high = "indianred2",
  limits = c(-750,2250))+xlab("")+
 theme(axis.text.x = element_text(angle = 90, color = "black"),
       axis.ticks.x = element_blank()
       ) + ggtheme()

#ggsave(filename = "figs/irea_gfd.svg", width = 19, height = 10, units = "cm")
ggsave(filename = "figs/irea_gfd_no_filtration.png", width = 60, height = 15, units = "cm")


```


```{r}
irea_all %>% 
   filter(Diet == "gf" & Cluster %in% c("cd4","cd8","treg")) %>% 
  filter(padj < 0.05) %>% 
 ggplot(aes(x = fct_reorder(Cytokine, `Enrichment Score`, .fun='median'), y = `Enrichment Score`, fill=..y..)) + 
 stat_summary(fun=mean, geom="bar", position = position_dodge(0.9)) +
      theme_bw()  + scale_size_continuous(range=c(1,4), limits = c(0,20)) +
  geom_point(aes(size = -log(padj, base = 10), shape = Cluster)) + 
  scale_shape_manual(values = c(17,19,15))+
 scale_fill_gradient2(
    low = "dodgerblue3",
  mid = "grey88",
  high = "indianred2",
  limits = c(-750,2250))+xlab("")+
 theme(axis.text.x = element_text(angle = 90, color = "black"),
       axis.ticks.x = element_blank()
       ) + ggtheme() + coord_flip()

#ggsave(filename = "figs/irea_gfd2.svg", width = 19, height = 22, units = "cm")
#ggsave(filename = "figs/irea_gfd2.png", width = 19, height = 22, units = "cm")

```


```{r}
irea_all %>% 
   filter(Diet == "std" & Cluster %in% c("cd4","cd8","treg")) %>% 
  filter(padj < 0.05) %>% 
 ggplot(aes(x = fct_reorder(Cytokine, `Enrichment Score`, .fun='median', .desc = TRUE), y = `Enrichment Score`, fill=..y..)) + 
 #geom_bar(position='dodge', stat='summary', fun='mean', alpha = 0.7, width = 0.9, aes(fill = color_enrich$enrich)) +
  stat_summary(fun=mean, geom="bar", position = position_dodge(0.9)) +
      theme_bw()  + scale_size_continuous(range=c(1,4), limits = c(0,20)) +
  geom_point(aes(size = -log(padj, base = 10), shape = Cluster)) + 
  scale_size_continuous(range = c(1,4))+
  scale_shape_manual(values = c(17,19,15))+
   scale_fill_gradient2(
    low = "dodgerblue3",
  mid = "grey88",
  high = "indianred2",
  limits = c(-750,2250))+xlab("")+
 theme(axis.text.x = element_text(angle = 90, color = "black"),
       axis.ticks.x = element_blank()
       ) + ggtheme()

ggsave(filename = "figs/irea_std.svg", width = 21, height = 10, units = "cm")
ggsave(filename = "figs/irea_std.png", width = 21, height = 10, units = "cm")


```

```{r}
irea_all %>% 
   filter(Diet == "std" & Cluster %in% c("cd4","cd8","treg")) %>% 
  filter(padj < 0.05) %>% 
 ggplot(aes(x = fct_reorder(Cytokine, `Enrichment Score`, .fun='median'), y = `Enrichment Score`, fill=..y..)) + 
 #geom_bar(position='dodge', stat='summary', fun='mean', alpha = 0.7, width = 0.9, aes(fill = color_enrich$enrich)) +
  stat_summary(fun=mean, geom="bar", position = position_dodge(0.9)) +
      theme_bw()  + scale_size_continuous(range=c(1,4), limits = c(0,20)) +
  geom_point(aes(size = -log(padj, base = 10), shape = Cluster)) + 
  scale_size_continuous(range = c(1,4))+
  scale_shape_manual(values = c(17,19,15))+
   scale_fill_gradient2(
    low = "dodgerblue3",
  mid = "grey88",
  high = "indianred2",
  limits = c(-750,2250))+xlab("")+
 theme(axis.text.x = element_text(angle = 90, color = "black"),
       axis.ticks.x = element_blank()
       ) + ggtheme() + coord_flip()

ggsave(filename = "figs/irea_std2.svg", width = 19, height = 24, units = "cm")
ggsave(filename = "figs/irea_std2.png", width = 19, height = 24, units = "cm")


```


# FC STD GF
```{r}
Idents(funda_merge) <- funda_merge$Diet
funda_merge$Diet %>% table

fc <- FoldChange(funda_merge, ident.1 = "GFD", ident.2 = "Standard")

fc %>% arrange(avg_log2FC)

```


```{r}
genes_hansen_up <- c('Cd247','Zbtb32','Galnt16','Slc7a6','Mboat1','Chst2','Cep55','Dennd2d','Nebl','Gfi1',
'AI838599','Adamts6','Podnl1','Itih5','Cxcl13','Ms4a4b','Zbp1','Gimap8','Cd28','Bcl11b','Thy1','Tnfsf13b',
'Bcl2','BC018473','Cd69','H2-Oa','Parm1','Tmtc1','Ms4a6b','Rgs16','Lipc','Atp8b4','Socs1','Ikzf2','Clu',
'Lyz2','Spns3','Gbp4','Stc2','Trat1','Mpzl3','Tox','Cst7','AC154378.1','Sit1','Cacnb3','Upp1',
'Lurap1l','Cd5','Themis','B4galt6','Gpr176','Il5ra','Il12rb1','Mcoln3','Pdcd1','Olfr1372-ps1',
'Tnf','Havcr1','Gbp6','Cd96','Colq','Trbv13-2','Nudt17','Cd8a','Dnaaf3','Gm44103','2010016I18Rik',
'Nlrc5','Ctla4','S1pr3','Gm2682','Eno2','Kcng1','Tmem154','Cd27','Tnn','Il17rb',
'Slfn1','Il2ra','Cish','Plekhf1','Dkk2','Mreg','Pla2g2d','Ggt1','Phf24','Mcemp1','Asprv1','Slc36a2','Trac',
'Tgtp1','Bcat1','Adam11','Gm39321','Trbv16','Pappa','5830411N06Rik','Rap1gap','Lamc2','Gm38244','Bst1',
'9330175E14Rik','Gm12407','9530052E02Rik','Ccr4','Nyx','Irak1bp1','Serpina3i','Stmn2','Smpd5','Chdh','Kcne4',
'Izumo1r','Gm11707','Ccl19-ps3','Tnfsf11','Scel','Slc12a5','Art2a-ps','Gm15056','Il7','A430088P11Rik',
'Madcam1','Msc','CT571247.2','Tnfsf8','Gm37975','Gm5134','Cilp','Tnfrsf9','Tmem45a','Ccdc30','Sspo',
'B630019A10Rik','Iglc4','Il13ra2','Cacna1b','Serpina3n','Drc1','Ccl21c','Cxcl9','Dntt','Il33','Erp27','Slco5a1',
'Lyz1','Art2b','Flrt3','Unc5c','2610528A11Rik','Pigr','Vmn2r96','Cdcp1','Gm10851','Cdhr1','Osbpl10','Pla2g4f',
'Fxyd2','Slc7a11','Gm12185','Rab26','Ryr3','Spock3','Kcnf1','Gxylt2','Ccl19','Gm48878','Gm16685','Slc10a6',
'A730020M07Rik','Col28a1','Gm13546','Gm47792','Dsc2','H2-M2','Nrg1','Grem1','Tmem150c','Thbs4','Mpped1','Trbv29','Dsc3',
'Serpina10','Slc6a7','Nrn1','Sostdc1','Fat3','2610035D17Rik','Gm10591','Aire','Gm33858','Gpnmb','Tarm1','Fbxo40',
'B430119L08Rik','Gbp10','Ccr8','Itgb8','Galnt6os','Car12','Tmprss11d','Hc','Insm1','Ighv12-3')
```

```{r}


fc.df  <- fc  %>%  arrange(desc(avg_log2FC)) %>% 
  rownames_to_column("gene") %>% 
  dplyr::select(gene, avg_log2FC)
ranks<- deframe(fc.df)

library(fgsea)

plotEnrichment(genes_hansen_up,
               ranks) + labs(title="Hansen up in Ctrl") + ylim(-0.7,0.7) + ggtheme() 
```


# Treg DE genes

```{r}

treg <- subset(funda_merge, annotations_coarse == "CD4 Treg")

DimPlot(treg)

mrk_de_treg <- FindAllMarkers(treg, only.pos = TRUE)
mrk_de_treg


avgexp = AverageExpression(treg, features = mrk_de_treg$gene,
                           return.seurat = F, group.by = "MouseID", 
                          assay = "RNA")

avgexp$RNA

options(repr.plot.width = 9.5, repr.plot.height = 2.5)
pheatmap(t(avgexp$RNA), main = "", 
         scale = "column", cluster_cols = F, cluster_rows = F,
        color=colorRampPalette(c("dodgerblue", "grey95", "indianred2"))(50), 
         border_color = "white",
                  fontsize = 12)


avgexp = AverageExpression(treg, features = c("Itgae","Klrg1","Tcf7","Foxp3","Tigit","Ctla4","Mki67"),
                           return.seurat = F, group.by = "MouseID", 
                          assay = "RNA")

avgexp$RNA

options(repr.plot.width = 9.5, repr.plot.height = 2.5)
pheatmap(t(avgexp$RNA), main = "", 
         scale = "column", cluster_cols = F, cluster_rows = F,
        color=colorRampPalette(c("dodgerblue", "grey95", "indianred2"))(50), 
         border_color = "white",
                  fontsize = 12)
```


# Clonal expansion
## Clone abundance

```{r}
metadata_6 <- funda_merge@meta.data
rownames(metadata_6) <- colnames(funda_merge)

metadata_6 <- metadata_6 %>% mutate(
  clone_nt = paste(cdr3_B_nt,cdr3_A1_nt,cdr3_A2_nt),
  clone_aa = paste("CDR3b",cdr3_B,"CDR3a",cdr3_A1)
)

clone_table <- metadata_6 %>%
  dplyr::group_by(factor(clone_aa)) %>%
          dplyr::summarize(n = n(), sum = sum()) %>%
    arrange(desc(n))

clone_table
```


```{r}
metadata_6$test <- 0
metadata_6 <- metadata_6 %>% group_by(test, clone_aa) %>% mutate(clone_abundance = as.numeric(n()))


metadata_6 <- as.data.frame(metadata_6 %>% mutate(clone_abundance = as.numeric(ifelse(clone_abundance==1056,NA_integer_,clone_abundance))) %>%
  mutate(log_clone_abundance = log(clone_abundance, base = 2)))
rownames(metadata_6) <- colnames(funda_merge)

funda_merge@meta.data <- metadata_6
rownames(funda_merge@meta.data) <- colnames(funda_merge)

funda_merge <- AddMetaData(funda_merge, as.numeric(metadata_6$clone_abundance), "clone_abundance")
rownames(funda_merge@meta.data) <- colnames(funda_merge)

FeaturePlot(funda_merge, reduction = "umap", features = "log_clone_abundance")

FeaturePlot(subset(funda_merge, clone_abundance >=1 & clone_abundance < 1056), reduction = "umap", features = "log_clone_abundance", cols = c("lightblue","firebrick")) + theme_classic() + theme(plot.title = element_text(hjust = 0.5)) + ggtitle("Log2 clone abundance")

#ggsave("fig/Log2CloneAbundance.png", width = 13, height = 8, units = "cm", dpi = 120)
#ggsave("fig/Log2CloneAbundance.svg", width = 13, height = 8, units = "cm", dpi = 120)

# Clone abundance bar graph
metadata_6 <- metadata_6 %>%
  mutate( clone_abundance_group = case_when(clone_abundance==5 ~ "5",
                                            clone_abundance==4 ~ "4",
                                            clone_abundance==3 ~ "3",
                                            clone_abundance==2 ~ "2",
                                            clone_abundance==1 ~ "1",
                                            TRUE ~ "0"
                                            ),
          clone_abundance_isNA = if_else(clone_abundance %in% c(1:5), "1", "0")) 

metadata_6 %>% 
  ggplot(aes(x = annotations_fine)) + 
  geom_bar(aes(fill = factor(clone_abundance_group)), position = "fill") + 
  coord_flip() +
  scale_fill_brewer(palette = "Blues") + xlab("Frequency")+
  ylab("") + 
  theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5), 
        legend.position="right", 
        panel.border = element_blank(), 
        legend.title = element_blank()) + 
  ggtitle("Clone abundance in clusters")

library(RColorBrewer)
brewer.pal(n, name)
```


```{r}
## TCR detection

metadata_6 %>% 
  ggplot(aes(x = factor(annotations_coarse, levels = rev(levels(factor(metadata_6$annotations_coarse)))))) + 
  geom_bar(aes(fill = factor(clone_abundance_isNA)), position = "fill") + 
  coord_flip() +
  scale_fill_manual(values = c("grey80", "palegreen3")) + xlab("Frequency")+
  ylab("") + 
  theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5), 
        legend.position="right", 
        panel.border = element_blank(), 
        legend.title = element_blank()) + 
  ggtitle("Clone abundance in clusters")
```


```{r}
### Clonal abundance, axis up to 30%
 metadata_6 %>% 
  filter(clone_abundance_isNA == 1) %>% 
  ggplot(aes(x = annotations_coarse)) + 
  geom_bar(aes(fill = factor(clone_abundance_group)), position = "fill") + 
   coord_flip() +
  scale_fill_brewer(palette = "Blues") + xlab("Frequency")+
  xlab("Frequency") +
  ylab("") + 
  theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5), legend.position="right", panel.border = element_blank(), legend.title = element_blank()) + 
  ggtitle("Clone abundance in clusters") + 
  ggtheme()

# ggsave("fig/Log2CloneAbundance_bar.png", width = 18, height = 10, units = "cm", dpi = 120)
# ggsave("fig/Log2CloneAbundance_bar.svg", width = 18, height = 10, units = "cm", dpi = 120)

```

```{r}

### Clonal abundance, axis up to 30%
 metadata_6 %>% 
    ggplot(aes(x = factor(annotations_coarse, levels = rev(levels(factor(annotations_coarse))[c(1,6,8,3,2,7,12,9,4,11,5,10)])))) + 
  geom_bar(aes(fill = factor(clone_abundance_group)), position = "fill") + 
   coord_flip() +
  scale_fill_brewer(palette = "Blues") + xlab("Frequency")+
  xlab("Frequency") +
  ylab("") + 
  theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5), legend.position="right", panel.border = element_blank(), legend.title = element_blank()) + 
  ggtitle("Clone abundance in clusters") + 
  ggtheme()

 ggsave("figv06/Log2CloneAbundance_bar_v06.png", width = 18, height = 10, units = "cm", dpi = 120)
 ggsave("figv06/Log2CloneAbundance_bar_v06.svg", width = 18, height = 10, units = "cm", dpi = 120)

```


```{r fig.width=30, fig.height=20}
plot_list <- list()

for(i in 2:16){
 
plot_list[[i]] <-  DimPlot(funda_merge, 
        cells.highlight = colnames(funda_merge)[which(funda_merge$clone_aa == pull(clone_table, `factor(clone_aa)`)[i])],
        sizes.highlight = 2) + NoLegend() + ggtitle(str_replace(string = pull(clone_table, `factor(clone_aa)`)[i], pattern = "CDR3a", replacement = "\nCDR3a"))
  }

cowplot::plot_grid(plotlist = plot_list[2:16], ncol = 5)
#ggsave("fig/tcr_funda.png", width = 35, height = 60, units = "cm")

```

## Clone abundance = boxplot 
```{r}

df4 <-metadata_6 %>% mutate(clon_expanded = if_else(clone_abundance_group == 1 ,"Unique clone","2+ clones")) %>% group_by(MouseID, clon_expanded) %>% 
  summarise(n = n()) %>%
  mutate(freq = n / sum(n)) 

df4 %>% group_by(MouseID) %>% summarise(sum = sum(freq))

    df4 %>%
      mutate(Diet = if_else(grepl(MouseID, pattern = "1|2|3|4|5"), "GFD", "STD")) %>% 
      filter(!is.na(clon_expanded)) %>% 
     ggplot(aes(x = Diet, y = freq*100)) + 
 geom_boxplot(outlier.shape = NA) +
 geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
      geom_jitter(position=position_jitter(0.2), size = 2, aes(color = Diet)) + 
      scale_color_manual(values= c("firebrick","lightskyblue3")) + 
      theme_minimal() + 
      facet_wrap(~clon_expanded, scales = "free", ncol = 2)  + ylim(c(0,NA))
    
  
```


## Clone abundance = boxplot2
```{r, fig.width=12, fig.height=8}

df4 <-metadata_6 %>% mutate(clon_expanded = if_else(clone_abundance_group == 1 ,"Unique clone","2+ clones")) %>% mutate(MouseID_cluster = paste(MouseID, seurat_clusters, sep = "-")) %>% group_by(MouseID_cluster, clon_expanded) %>% 
  summarise(n = n()) %>%
  mutate(freq = n / sum(n)) 

df4

df4 %>% group_by(MouseID_cluster) %>% summarise(sum = sum(freq))

    df4 %>%
      mutate(Diet = if_else(grepl(MouseID_cluster, pattern = "_1|_2|_3|_4|_5"), "GFD", "STD")) %>% 
      separate(MouseID_cluster, into = c("MouseID", "seurat_clusters"), sep = "-") %>% 
       filter(!is.na(clon_expanded)) %>% 
           filter(clon_expanded == "2+ clones") %>% 
     ggplot(aes(x = Diet, y = freq*100)) + 
 geom_boxplot(outlier.shape = NA) +
 geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
      geom_jitter(position=position_jitter(0.2), size = 2, aes(color = Diet)) + 
      scale_color_manual(values= c("firebrick","lightskyblue3")) + 
      theme_minimal() + 
      facet_wrap(~factor(seurat_clusters, levels = c(0:9,11,17)), scales = "free", ncol = 7)  + ylim(c(0,NA))
    
  
```



Clones in tissues

```{r}

## top 10 clones tissues 

top_10_clones <- pull(clone_table, `factor(clone_aa)`)[2:11]

top_10_clones %>% 
  kbl() %>% 
  kable_styling(font_size = 11,
                bootstrap_options = c("striped", "hover", 
                                      "condensed", "responsive")) %>% 
  scroll_box(width = "100%", height = "400px")

metadata_7 <- metadata_6 %>% filter(clone_aa %in% top_10_clones) %>% group_by(clone_aa, Tissue) %>% tally %>% pivot_wider(names_from = "Tissue", values_from = n, values_fill = 0) %>% arrange(desc(Spleen))

```


# Reclustering unconventional and proliferating cells

## Subsetting

```{r}
DimPlot(funda_merge, label = T)

unconv_prolif <- subset(funda_merge, seurat_clusters %in% c(7,9,10,16,17))

 unconv_prolif <- NormalizeData(object = unconv_prolif)
  unconv_prolif <- ScaleData(unconv_prolif, verbose = FALSE)
  unconv_prolif <- FindVariableFeatures(unconv_prolif, selection.method = "vst", nfeatures = 1000, 
                                       verbose = FALSE)
  unconv_prolif <- RunPCA(unconv_prolif, npcs = 10, verbose = FALSE)
  unconv_prolif <- RunUMAP(unconv_prolif, reduction = "pca", dims = 1:10)
  unconv_prolif <- FindNeighbors(unconv_prolif, dims = 1:10)
  unconv_prolif <- FindClusters(unconv_prolif, resolution = 0.9)
  DimPlot(unconv_prolif, label = T)
  

  DimPlot(unconv_prolif, group.by = "Tissue") 
  DimPlot(unconv_prolif, group.by = "Diet") 
  

FeaturePlot(unconv_prolif, features = c("Cd4", "Cd8a", "Mki67", "Ncr1"))
FeaturePlot(unconv_prolif, features = c("Foxp3", "Il2ra", "Trdc", "Trgv2"))
FeaturePlot(unconv_prolif, features = c("Gata3", "Tbx21", "Ifng", "Bcl6"))
FeaturePlot(unconv_prolif, features = c("Ikzf2", "Klrk1", "Klrd1", "Il7r"))

```

```{r fig.width=18, fig.height=10}

unconv_prolif <- reduced_annot(unconv_prolif, "Immgen_annot_single", n_cells_to_other = 20)

unconv_prolif$Immgen_annot_single_red %>% table

DimPlot(unconv_prolif, group.by = "Immgen_annot_single_red", label = T, repel = T) 
 
#ggsave("fig/unconventional_dimplot.png", width = 35, height = 25, units = "cm")
```

## Unconventional proliferating annotations

```{r fig.width=7, fig.height=4}
  DimPlot(unconv_prolif, label = T)

md <- unconv_prolif@meta.data

md <- md %>% mutate(annotations_fine = recode(seurat_clusters, 
                                                     "0" = "00 NK cells 1",
                                                     "1" = "01 CD4 naive",
                                                     "2" = "02 Treg 1",
                                                     "3" = "03 Tgd 1",
                                                     "4" = "04 Proliferating",
                                                     "5" = "05 NKT / Th1",
                                                     "6" = "06 Treg 2",
                                                     "7" = "07 CD8 naive",
                                                     "8" = "08 NK and Tgd 3",
                                                     "9" = "09 Tgd 2",
                                                     "10" = "10 NK cells 2",
                                                     "11" = "11 NKT / Th17"))

unconv_prolif@meta.data <- md
DimPlot(unconv_prolif, group.by = "annotations_fine")
```

## Markers

```{r eval=TRUE, include=TRUE}
markers_prolif <- FindAllMarkers(unconv_prolif, min.diff.pct = 0.05, logfc.threshold = log(1.5))

markers_prolif <- rank_score_func(markers_prolif)
markers_prolif %>% filter(avg_log2FC >0) %>% group_by(cluster) %>% slice_max(order_by = score, n = 20) %>% 
  kbl() %>% 
  kable_styling(font_size = 11,
                bootstrap_options = c("striped", "hover", 
                                      "condensed", "responsive")) %>% 
  scroll_box(width = "100%", height = "400px")


```

```{r fig.width=30, fig.height=23}
FeaturePlot(unconv_prolif, features = c("Ncr1","Gzma","Fcer1g","Prf1","Klre1",
                                        "Cd4", "Cd40lg","S1pr1","Ccr7", "Il6ra",
                                        "Foxp3","Il1rl1","Ctla4","Tigit","Icos",
                                        "Trgv2","Trdc","Ccr9","Itgae","Tmem108",
                                        "Cxcr6","Il4","Zbtb16","Xcl1","Fasl",
                                        "Cd200","Izumo1r","Tnfsf11","Tnfsf8","Pdcd1"),
            ncol = 5)
#ggsave("fig/unconv_markers1.png", width = 30, height = 20)

FeaturePlot(unconv_prolif, features = c("Gzmk","Ccr10","Cx3cr1","S100a4","Lgals3",
                                        "Kcnj8", "Gzmm","Ncr1","Cd7", "Xcl1",
                                        "Hck","Prdm16","Tcrg-V7","Cxxc5","Klri1",
                                        "Tnni1","Klrb1a","Scimp","Klrb1f","Klrb1f",
                                        "Blk","Lingo4","Il17re","Rorc","Il23r"),
            ncol = 5)
#ggsave("fig/unconv_markers2.png", width = 30, height = 20)

```

Good vs bad NK cells

```{r}

nk_cell_diff <- FindMarkers(unconv_prolif, ident.1 = 0, ident.2 = 10)
#nk_cell_diff

```


Boxplots

```{r}
seurat_meta_data <- unconv_prolif@meta.data

# Create grouped dataframe, calculate the frequencies of clusters
df4 <- seurat_meta_data %>% 
  group_by(MouseID, annotations_fine) %>% 
  summarise(n = n()) %>% 
  unique() %>% 
  mutate(freq = n / sum(n)) 

# Control - all sums should be one
# df4 %>% group_by(MouseID) %>% summarise(sum = sum(freq))

# As we've lost non-grouping variables, let's join them back
md_to_join <- seurat_meta_data %>% 
  dplyr::select(MouseID, Diet, annotations_fine) %>% 
  unique()

df4  <- left_join(df4, md_to_join)

# The final plot

df4 %>% 
  ggplot(aes(x = Diet, y = freq)) + # you can change the x to whatever variable you're interested in
  geom_boxplot(outlier.shape = NA) + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(position=position_jitter(0.2), size = 2, aes(color = Diet)) + 
  facet_wrap(~annotations_fine, scales = "free", ncol = 5) +
  ylab("Frequency") +
  xlab("Condition") +
  theme_classic() +
  theme(strip.background = element_blank(), panel.grid = element_blank()) + 
  theme(axis.text.x = element_text(angle = 90))


# #ggsave("fig/clusters_boxplots_unconventional.png", width = 20, height = 15, units = "cm")

```


## Unconventional clones 

```{r}
FeaturePlot(unconv_prolif, reduction = "umap", features = "log_clone_abundance")

FeaturePlot(subset(unconv_prolif, clone_abundance >=1 & clone_abundance < 1056), reduction = "umap", features = "log_clone_abundance", cols = c("lightblue","firebrick")) + theme_classic() + theme(plot.title = element_text(hjust = 0.5)) + ggtitle("Log2 clone abundance")

```


```{r fig.width=20, fig.height=35}
metadata_6 <- unconv_prolif@meta.data
rownames(metadata_6) <- colnames(unconv_prolif)

metadata_6 <- metadata_6 %>% mutate(
  clone_nt = paste(cdr3_B_nt,cdr3_A1_nt,cdr3_A2_nt),
  clone_aa = paste("CDR3b",cdr3_B,"CDR3a",cdr3_A1)
)

clone_table_prolif <- metadata_6 %>%
  dplyr::group_by(factor(clone_aa)) %>%
          dplyr::summarize(n = n(), sum = sum()) %>%
    arrange(desc(n))

clone_table_prolif %>% 
  kbl() %>% 
  kable_styling(font_size = 11,
                bootstrap_options = c("striped", "hover", 
                                      "condensed", "responsive")) %>% 
  scroll_box(width = "100%", height = "400px")
```


```{r fig.width=20, fig.height=35}
plot_list <- list()

for(i in 2:15){
 
plot_list[[i]] <-  DimPlot(unconv_prolif, split.by = "Diet",
        cells.highlight = colnames(unconv_prolif)[which(unconv_prolif$clone_aa == pull(clone_table_prolif, `factor(clone_aa)`)[i])],
        sizes.highlight = 2) + NoLegend() + ggtitle(str_replace(string = pull(clone_table_prolif, `factor(clone_aa)`)[i], pattern = "CDR3a", replacement = "\nCDR3a"))
}

cowplot::plot_grid(plotlist = plot_list[2:15], ncol = 2)
```

# Show Egr and early activated

```{r fig.height = 4, fig.width = 5}
funda_merge$c14 <- funda_merge$seurat_clusters %in% c(14)
DimPlot(funda_merge, group.by = "c14", cols = c("grey78","#b532b2ff"))

funda_merge$c18 <- funda_merge$seurat_clusters %in% c(18)
DimPlot(funda_merge, group.by = "c18", cols = c("grey78","#b532b2ff"))

```


# Treg reclustering

```{r}
DimPlot(funda_merge, label = T, group.by = "seurat_clusters")

tregs_tfh <- subset(funda_merge, seurat_clusters %in% c(7,8,11))

 tregs_tfh <- NormalizeData(object = tregs_tfh)
  tregs_tfh <- ScaleData(tregs_tfh, verbose = FALSE)
  tregs_tfh <- FindVariableFeatures(tregs_tfh, selection.method = "vst", nfeatures = 1000, 
                                       verbose = FALSE)
  tregs_tfh <- RunPCA(tregs_tfh, npcs = 10, verbose = FALSE)
  tregs_tfh <- RunUMAP(tregs_tfh, reduction = "pca", dims = 1:10)
  tregs_tfh <- FindNeighbors(tregs_tfh, dims = 1:10)
  tregs_tfh <- FindClusters(tregs_tfh, resolution = 0.4)
  DimPlot(tregs_tfh, label = T)
  

  DimPlot(tregs_tfh, group.by = "Tissue") 
  DimPlot(tregs_tfh, group.by = "Diet") 
  


FeaturePlot(tregs_tfh, features = c("Mki67", "Cxcr6", "Foxp3", "Ikzf2"))
FeaturePlot(tregs_tfh, features = c("Cd4", "Cd8a", "Mki67", "Rorc"))
FeaturePlot(tregs_tfh, features = c("Cd3d", "Il1r1", "Cd160", "Ncr1"))
FeaturePlot(tregs_tfh, features = c("Foxp3", "Il2ra", "Trdc", "Trgv2"))
FeaturePlot(tregs_tfh, features = c("Gata3", "Tbx21", "Ifng", "Bcl6"))
FeaturePlot(tregs_tfh, features = c("Ikzf2", "Bcl6", "Klrd1", "Il7r"))
FeaturePlot(tregs_tfh, features = c("Tnfrsf18", "Icos", "Itgae", "Pdcd1"))



FeaturePlot(tregs_tfh, features = c("Cd8a", "Foxp3"))
VlnPlot(tregs_tfh, features = c("Cd8a", "Foxp3"), group.by = )

tregs_tfh$annotations_fine %>% table

```

# Fig plot

```{r fig.height = 4, fig.width = 5}
funda_merge$treg <- funda_merge$seurat_clusters %in% c(7,8,11)
DimPlot(funda_merge, group.by = "treg", cols = c("grey78","#fd7a00ff"))

DimPlot(tregs_tfh, cols = c("#e3e34dff", "#5a9fdeff","#d5a43fff",
                            "#6faea9ff","#d2723dff","#c94eb8ff"), pt.size = 6, raster = T,
        raster.dpi = c(900,900)) + ggtheme()
#ggsave("fig/treg_dimplot.svg", width = 8, height = 5.5, units = "cm")



```

## Markers
```{r}
markers_tregs <- FindAllMarkers(tregs_tfh, min.diff.pct = 0.05, logfc.threshold = log(1.5))

markers_tregs <- rank_score_func(markers_tregs)


markers_tregs %>% filter(avg_log2FC >0) %>% group_by(cluster) %>% slice_max(order_by = score, n = 20)


markers_tregs %>% filter(avg_log2FC >0) %>% group_by(cluster) %>% slice_max(order_by = score, n = 20) %>% 
  kbl() %>% 
  kable_styling(font_size = 11,
                bootstrap_options = c("striped", "hover", 
                                      "condensed", "responsive")) %>% 
  scroll_box(width = "100%", height = "400px")


```

## Markers
```{r}
table(Idents(tregs_tfh))
markers_tregs <- FindAllMarkers(tregs_tfh, only.pos = T)

markers_tregs
write.csv(markers_tregs, "240806_markers_treg.csv")
```


## Heatmap Treg markers

```{r fig.width = 9.5, fig.height = 2.5}
markers  <- str_to_sentence(rev(c("IKZF2","IL2RA","FOXP3","KLRG1",
                                  "CCR7","SELL",  "TCF7", "CD44", "IL7R",
                                  "CTLA4","IL2RA","IL10",
                                  "BCL6","CXCR5","TBX21",
                                  "Cd40lg","Itgae","Tigit","Il18r","S100a6","Il1rl1","S100a4","Gzmb",
                                  "Egr1","Tnfrsf9","Egr2","Cd200","Irf4","Lag3",
                                  "Satb1","Il21","Pdcd1",
                                  "MKI67","IFNG","Ccr10"
                  )))


avgexp = AverageExpression(tregs_tfh, features = markers,
                           return.seurat = F, group.by = "seurat_clusters", 
                          assay = "RNA")

pheatmap(t(avgexp$RNA), main = "",
         scale = "column", cluster_cols = T, cluster_rows = T,
        color=colorRampPalette(c("dodgerblue", "grey95", "indianred2"))(50),
         border_color = "white",
                  fontsize = 12)


# pheatmap(t(avgexp$RNA[,c(1,6,3,2,7,8,4,10,5,9)]), main = "", 
#          scale = "column", cluster_cols = F, cluster_rows = F,
#         color=colorRampPalette(c("dodgerblue", "grey95", "indianred2"))(50), 
#          border_color = "white",
#                   fontsize = 12)


markers_order  <- rev(c("Ctla4","Il2ra","Foxp3","Ikzf2","Sell","Ccr7","Satb1","Lef1","Tcf7",
                    
                    "Ccr10","Il7r","Tbx21","Cd44",
                    "Ifng","Irf4","Tnfrsf9","Egr1",
                    "S100a6","S100a4","Gzmb","Klrg1",
                    "Icos","Itgae",
                    "Bcl6","Pdcd1","Il21","Cxcr5"))

avgexp = AverageExpression(tregs_tfh, features = markers_order,
                           return.seurat = F, group.by = "seurat_clusters", 
                          assay = "RNA")
pheatmap(t(avgexp$RNA)[c(1,4,2,3,5,6),], main = "",
         scale = "column", cluster_cols = F, cluster_rows = F,
        color=colorRampPalette(c("dodgerblue", "grey95", "indianred2"))(50),
         border_color = "white",
                  fontsize = 12)

pheatmap::pheatmap(t(avgexp$RNA)[c(1,4,2,3,5,6),], main = "",
         scale = "column", cluster_cols = F, cluster_rows = F,
        color=colorRampPalette(c("dodgerblue", "grey95", "indianred2"))(50),
         border_color = "white",
                  fontsize = 12, 
        filename = "plots/heatmap_treg_v03.pdf",
         width = 6, height = 2.2,)
```


```{r}
markers  <- str_to_sentence(rev(c("ICOS","ITGAE","IKZF2","IL2RA","FOXP3","KLRG1",
                                  "CCR7","SELL",  "TCF7", "CD44", "IL7R",
                                  "CTLA4","IL2RA","IL10",
                                  "BCL6","CXCR5","TBX21",
                                  "Cd40lg","Itgae","Tigit","Il18r","S100a6","Il1rl1","S100a4","Gzmb",
                                  "Egr1","Tnfrsf9","Egr2","Cd200","Irf4","Lag3",
                                  "Satb1","Il21","Pdcd1",
                                  "MKI67","IFNG","Ccr10"
                  )))


avgexp = AverageExpression(tregs_tfh, features = markers,
                           return.seurat = F, group.by = "seurat_clusters", 
                          assay = "RNA")

pheatmap(t(avgexp$RNA)[c(1,4,2,3,5,6),], main = "",
         scale = "column", cluster_cols = F, cluster_rows = F,
        color=colorRampPalette(c("dodgerblue", "grey95", "indianred2"))(50),
         border_color = "white",
                  fontsize = 12)
```

## Treg proliferating annotations

```{r fig.width=9, fig.height=3}
DimPlot(tregs_tfh, label = T)

md <- tregs_tfh@meta.data

md <- md %>% mutate(annotations_fine = recode(seurat_clusters, 
                                                     "0" = "1 Treg naive",
                                                     "1" = "5 Treg memory",
                                                     "2" = "2 Treg early activated",
                                                     "3" = "4 Treg stem ",
                                                     "4" = "3 Treg cytotoxic",
                                                     "5" = "5 Tfh"))

tregs_tfh@meta.data <- md
DimPlot(tregs_tfh, group.by = "annotations_fine")
```

## Treg Boxplots

```{r fig.width = 12, fig.height = 4.5}
seurat_meta_data <- tregs_tfh@meta.data
seurat_meta_data$MouseID2 <- str_replace_all(string = seurat_meta_data$MouseID, 
                                            pattern = "Mouse_", 
                                            replacement = "")
# Create grouped dataframe, calculate the frequencies of clusters
df4 <- seurat_meta_data %>% 
  group_by(MouseID, seurat_clusters) %>% 
  summarise(n = n()) %>% 
  unique() %>% 
  mutate(freq = n / sum(n)) 

# Control - all sums should be one
# df4 %>% group_by(MouseID) %>% summarise(sum = sum(freq))

# As we've lost non-grouping variables, let's join them back
md_to_join <- seurat_meta_data %>% 
  dplyr::select(MouseID, Diet, seurat_clusters, MouseID2) %>% 
  unique()

df4  <- left_join(df4, md_to_join)

# The final plot

df4 %>% 
  ggplot(aes(x = Diet, y = freq)) + # you can change the x to whatever variable you're interested in
  geom_boxplot(outlier.shape = NA) + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_point(shape = 21, size = 6, aes(fill = Diet), ) +
  geom_text(aes(label = MouseID2, x = Diet, y = freq), size = 4) +
  facet_wrap(~seurat_clusters, scales = "free", ncol = 6) +
  ylab("Frequency") +
  xlab("Condition") +
  theme_classic() +
  scale_fill_manual(values = c("#bfdbf7ff", "#f6b0b0ff")) +
  theme(strip.background = element_blank(), panel.grid = element_blank()) + 
  theme(axis.text.x = element_text(angle = 90))

 #ggsave("fig/clusters_boxplots_tregs.png", width = 20, height = 5, units = "cm")

```

```{r}

seurat_meta_data <- tregs_tfh@meta.data
seurat_meta_data$MouseID2 <- str_replace_all(string = seurat_meta_data$MouseID, 
                                            pattern = "Mouse_", 
                                            replacement = "")
# Create grouped dataframe, calculate the frequencies of clusters
df4 <- seurat_meta_data %>% 
  group_by(MouseID, annotations_fine) %>% 
  summarise(n = n()) %>% 
  unique() %>% 
  mutate(freq = n / sum(n)) 

# Control - all sums should be one
# df4 %>% group_by(MouseID) %>% summarise(sum = sum(freq))

# As we've lost non-grouping variables, let's join them back
md_to_join <- seurat_meta_data %>% 
  dplyr::select(MouseID, Diet, annotations_fine, MouseID2) %>% 
  unique()

df4  <- left_join(df4, md_to_join)

df4 %>% 
  ggplot(aes(x = Diet, y = freq*100)) + # you can change the x to whatever variable you're interested in
  geom_boxplot(outlier.shape = NA, aes(fill = Diet), alpha = 0.7) + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(width = 0.1, height = 0.1, shape = 21,
              size = 4, aes(fill = Diet)) + # in aes, you can also use shape or fill (for the shapes that allow it)
  facet_wrap(~annotations_fine, scales = "free", ncol = 6) +
  scale_fill_manual(values = c("dodgerblue","indianred2")) + 
  scale_color_manual(values = c("dodgerblue","indianred2")) + 
  ylab("Frequency") +
  xlab("Condition") +
  theme_classic() +
  ylim(0,NA)+
  theme(strip.background = element_blank(), panel.grid = element_blank()) + 
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) + ggtitle("Spleen scRNAseq") +
  ggpubr::stat_compare_means(label.x = 1.2, label.y.npc = "top", size = 5, vjust = 0.3, label = "p.format") + 
  ggtheme()

#ggsave("figs/treg_quant.svg", width = 32, height = 9, units = "cm")
```



### Quantification tissue

```{r}
seurat_meta_data <- tregs_tfh@meta.data
# Create grouped dataframe, calculate the frequencies of clusters
df4 <- seurat_meta_data %>% 
  mutate(MouseID_tissue = paste(MouseID, Tissue)) %>% 
  group_by(MouseID_tissue, annotations_fine) %>% 
  summarise(n = n()) %>% 
  unique() %>% 
  mutate(freq = n / sum(n)) 

# Control - all sums should be one
# df4 %>% group_by(MouseID_tissue) %>% summarise(sum = sum(freq))

# As we've lost non-grouping variables, let's join them back
md_to_join <- seurat_meta_data %>% 
  dplyr::select(MouseID, Diet, Tissue) %>% 
  unique() %>% 
  mutate(MouseID_tissue = paste(MouseID, Tissue))
df4  <- left_join(df4, md_to_join)

# The final plot
df4 %>% 
  ggplot(aes(x = factor(Tissue, labels = c("pLN", "SPL")), y = freq*100)) + # you can change the x to whatever variable you're interested in
  geom_boxplot(outlier.shape = NA, aes(fill = Tissue), alpha = 0.5) +
 geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(shape=21, size = 3, position=position_jitter(width = 0.1, height = 0), 
              aes(fill = Tissue)) + theme_classic() + xlab("Group") +  
  theme_classic() +
  theme(strip.background = element_blank(), panel.grid = element_blank()) + 
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) + ggtitle("Treg scRNAseq by tissue") +
  ggpubr::stat_compare_means(label.x = 1.2, label.y.npc = "top", size = 5, vjust = 0.3, label = "p.format") + 
  ggtheme() +
  scale_color_manual(values = c("#5fd3bcff","#4a6ca3ff")) + 
  scale_fill_manual(values = c("#5fd3bcff","#4a6ca3ff")) +
  facet_wrap(~annotations_fine, scales = "free", ncol = 6) +
  ylim(0,NA) +
  ylab("Frequency") +
  xlab("Condition") 

  #ggsave("figs/treg_quant_tissue.svg", width = 32, height = 9, units = "cm")

```

### Quantification Diet

```{r}
# Create grouped dataframe, calculate the frequencies of clusters
df4 <- seurat_meta_data %>% 
  mutate(MouseID_tissue = paste(MouseID, Tissue)) %>% 
  group_by(MouseID_tissue, annotations_fine) %>% 
  summarise(n = n()) %>% 
  unique() %>% 
  mutate(freq = n / sum(n)) 

# Control - all sums should be one
# df4 %>% group_by(MouseID_tissue) %>% summarise(sum = sum(freq))

# As we've lost non-grouping variables, let's join them back
md_to_join <- seurat_meta_data %>% 
  dplyr::select(MouseID, Diet, Tissue) %>% 
  unique() %>% 
  mutate(MouseID_tissue = paste(MouseID, Tissue))
df4  <- left_join(df4, md_to_join)

# The final plot
df4 %>% 
  dplyr::filter(Tissue == "Spleen") %>% 
  ggplot(aes(x = Diet, y = freq*100)) + # you can change the x to whatever variable you're interested in
  geom_boxplot(outlier.shape = NA, aes(fill = Diet), alpha = 0.5) +
 geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(shape=21, size = 3, position=position_jitter(width = 0.1, height = 0), 
              aes(fill = Diet)) + theme_classic() + xlab("Group") +  
  theme_classic() +
  theme(strip.background = element_blank(), panel.grid = element_blank()) + 
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) + ggtitle("Treg scRNAseq by diet, spleen") +
  ggpubr::stat_compare_means(label.x = 1.2, label.y.npc = "top", size = 5, vjust = 0.3, label = "p.format") + 
  ggtheme() +
  scale_fill_manual(values = c("dodgerblue","indianred2")) + 
  scale_color_manual(values = c("dodgerblue","indianred2")) + 
  facet_wrap(~annotations_fine, scales = "free", ncol = 6) +
  ylim(0,NA) +
  ylab("Frequency") +
  xlab("Condition") 

  #ggsave("figs/treg_quant_tissue.svg", width = 32, height = 9, units = "cm")

```

```{r}
df4 %>% 
  dplyr::filter(Tissue == "pLN") %>% 
  ggplot(aes(x = Diet, y = freq*100)) + # you can change the x to whatever variable you're interested in
  geom_boxplot(outlier.shape = NA, aes(fill = Diet), alpha = 0.5) +
 geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(shape=21, size = 3, position=position_jitter(width = 0.1, height = 0), 
              aes(fill = Diet)) + theme_classic() + xlab("Group") +  
  theme_classic() +
  theme(strip.background = element_blank(), panel.grid = element_blank()) + 
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) + ggtitle("Treg scRNAseq by diet, spleen") +
  ggpubr::stat_compare_means(label.x = 1.2, label.y.npc = "top", size = 5, vjust = 0.3, label = "p.format") + 
  ggtheme() +
  scale_fill_manual(values = c("dodgerblue","indianred2")) + 
  scale_color_manual(values = c("dodgerblue","indianred2")) + 
  facet_wrap(~annotations_fine, scales = "free", ncol = 6) +
  ylim(0,NA) +
  ylab("Frequency") +
  xlab("Condition") 

```


## Quantification in full dataset

```{r}
tregs_tfh$seurat_clusters %>% table
DimPlot(tregs_tfh)

tregs_tfh$annot_treg <- paste("Treg cl.",tregs_tfh$seurat_clusters)
tregs_tfh$barcode <- colnames(tregs_tfh)
```

```{r}
funda_merge$barcode <- colnames(funda_merge)
funda_merge@meta.data <- left_join(funda_merge@meta.data, dplyr::select(tregs_tfh@meta.data, barcode, annot_treg))

rownames(funda_merge@meta.data) <- colnames(funda_merge)
```

```{r}
DimPlot(tregs_tfh, label = T, label.size = 9)
DimPlot(funda_merge, group.by = "annot_treg")
```

```{r fig.width = 12, fig.height = 4.5}
seurat_meta_data <- funda_merge@meta.data
seurat_meta_data$MouseID2 <- paste(str_replace_all(string = seurat_meta_data$MouseID, 
                                            pattern = "Mouse_", 
                                            replacement = ""),seurat_meta_data$Tissue)
# Create grouped dataframe, calculate the frequencies of clusters
df4 <- seurat_meta_data %>% 
  group_by(MouseID, annot_treg) %>% 
  summarise(n = n()) %>% 
  unique() %>% 
  mutate(freq = n / sum(n)) 

# Control - all sums should be one
# df4 %>% group_by(MouseID) %>% summarise(sum = sum(freq))

# As we've lost non-grouping variables, let's join them back
md_to_join <- seurat_meta_data %>% 
  dplyr::select(MouseID, Diet, annot_treg, Tissue, MouseID2) %>% 
  unique()

df4  <- left_join(df4, md_to_join)

# The final plot

df4 %>% 
  ggplot(aes(x = Diet, y = freq*100)) + # you can change the x to whatever variable you're interested in
  geom_boxplot(outlier.shape = NA, aes(fill = Diet), alpha = 0.7) + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(width = 0.1, height = 0.1, shape = 21,
              size = 4, aes(fill = Diet)) + # in aes, you can also use shape or fill (for the shapes that allow it)
  facet_wrap(~annot_treg, scales = "free", ncol = 6) +
  scale_fill_manual(values = c("dodgerblue","indianred2")) + 
  scale_color_manual(values = c("dodgerblue","indianred2")) + 
  ylab("Frequency") +
  xlab("Condition") +
  theme_classic() +
  ylim(0,NA)+
  theme(strip.background = element_blank(), panel.grid = element_blank()) + 
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) + ggtitle("Both tissues scRNAseq") +
  ggpubr::stat_compare_means(label.x = 1.2, label.y.npc = "top", size = 5, vjust = 0.3, label = "p.format") + 
  ggtheme()

```


```{r}
DimPlot(tregs_tfh)

df4 %>% 
  dplyr::filter(!is.na(annot_treg) & Tissue == "pLN") %>% 
  ggplot(aes(x = Diet, y = freq*100)) + # you can change the x to whatever variable you're interested in
  geom_boxplot(outlier.shape = NA, aes(fill = Diet), alpha = 0.7) + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(width = 0.1, height = 0.1, shape = 21,
              size = 4, aes(fill = Diet)) + # in aes, you can also use shape or fill (for the shapes that allow it)
  facet_wrap(~factor(annot_treg, levels = levels(factor(df4$annot_treg))[c(1,3,5,4,2,6)]),
             scales = "free", ncol = 6) +
  scale_fill_manual(values = c("dodgerblue","indianred2")) + 
  scale_color_manual(values = c("dodgerblue","indianred2")) + 
  ylab("Frequency") +
  xlab("Condition") +
  theme_classic() +
  ylim(0,NA)+
  theme(strip.background = element_blank(), panel.grid = element_blank()) + 
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) + ggtitle("pLN scRNAseq") +
  ggpubr::stat_compare_means(label.x = 1.2, label.y.npc = "top", size = 5, vjust = 0.3, label = "p.format", method = "t.test") + 
  ggtheme()

  ggsave("figv06/treg_quant_diet_pLN_ttest.png", width = 32, height = 9, units = "cm")

```

```{r}
df4 %>% 
  dplyr::filter(!is.na(annot_treg) & Tissue == "Spleen") %>% 
  ggplot(aes(x = Diet, y = freq*100)) + # you can change the x to whatever variable you're interested in
  geom_boxplot(outlier.shape = NA, aes(fill = Diet), alpha = 0.7) + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(width = 0.1, height = 0.1, shape = 21,
              size = 4, aes(fill = Diet)) + # in aes, you can also use shape or fill (for the shapes that allow it)
  facet_wrap(~factor(annot_treg, levels = levels(factor(df4$annot_treg))[c(1,3,5,4,2,6)]),
             scales = "free", ncol = 6) +
  scale_fill_manual(values = c("dodgerblue","indianred2")) + 
  scale_color_manual(values = c("dodgerblue","indianred2")) + 
  ylab("Frequency") +
  xlab("Condition") +
  theme_classic() +
  ylim(0,NA)+
  theme(strip.background = element_blank(), panel.grid = element_blank()) + 
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) + ggtitle("SPL scRNAseq") +
  ggpubr::stat_compare_means(label.x = 1.2, label.y.npc = "top", size = 5, vjust = 0.3, label = "p.format", method = "t.test") + 
  ggtheme()

 # ggsave("figv06/treg_quant_diet_SPL_ttest.svg", width = 32, height = 9, units = "cm")

ggsave("figv06/treg_quant_diet_SPL_ttest.png", width = 32, height = 9, units = "cm")
```



# Treg FACS 

## Treg 6 clusters

```{r}
hm <- read_xlsx("../../../Project Funda NOD/FACS/Final_analysis/HighDim/240626_treg_6clusters_heatmap.xlsx")

hm_map <- hm %>% column_to_rownames("Cluster") %>% as.matrix()

hm_map

pheatmap(hm_map, main = "", 
         scale = "column", cluster_cols = T, cluster_rows = T,
        color=colorRampPalette(c("dodgerblue", "grey95", "indianred2"))(50), 
         border_color = "white",
                  fontsize = 12)

pheatmap(hm_map[c(4,3,2,1,6,5),c(5,6,1,2,4,3)], main = "", 
         scale = "column", cluster_cols = F, cluster_rows = F,
        color=colorRampPalette(c("dodgerblue", "grey95", "indianred2"))(50), 
         border_color = "white",
                  fontsize = 12, width = 4, height = 3)

pheatmap::pheatmap(hm_map[c(4,3,2,1,6,5),c(5,6,1,2,4,3)], main = "", 
         scale = "column", cluster_cols = F, cluster_rows = F,
        color=colorRampPalette(c("dodgerblue", "grey95", "indianred2"))(50), 
         border_color = "white",
                  fontsize = 12, width = 4, height = 3,
         filename = "figs/heatmap_cluster_FACS_Treg.pdf")

getwd()
```

```{r}
high_dim <- read_excel("G:/48_lab/Project Funda NOD/FACS/Final_analysis/HighDim/240626_treg_6clusters.xlsx")


high_dim <- t(high_dim) %>% as.data.frame() %>% rownames_to_column("sample")

high_dim
high_dim <- high_dim[2:217,]
high_dim
colnames(high_dim) <- c("population", "value")

high_dim
high_dim$population <- sub(pattern = "X-CE_Set_02b6dd_", 
                           replacement = "cluster",
                            x = high_dim$population)
high_dim$population <- sub(pattern = "\\| Freq. of Parent", 
                           replacement = "",
                            x = high_dim$population)

high_dim


high_dim <- high_dim %>% separate(population, into = c("sample","cluster"), 
                                  sep = "/", remove = F) %>% mutate(
                                    value = sub(value, pattern = ",", replacement = ".")
                                  ) %>% mutate(value = as.numeric(value))

high_dim <- high_dim %>% filter(!is.na(cluster))


high_dim
```


```{r fig.width=12, fig.height=4}
md <- read_excel("G:/48_lab/Project Funda NOD/FACS/Final_analysis/HighDim/240626_treg_14clusters_metadata.xlsx")

high_dim2 <- left_join(high_dim, md)

high_dim2

high_dim2 %>% 
  dplyr::filter(Tissue == "SPL") %>% 
  ggplot(aes(x = Diet, y = value)) + # you can change the x to whatever variable you're interested in
  geom_boxplot(outlier.shape = NA, aes(fill = Diet), alpha = 0.7) + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(width = 0.1, height = 0.1, shape = 21,
              size = 4, aes(fill = Diet)) + # in aes, you can also use shape or fill (for the shapes that allow it)
  facet_wrap(~cluster, scales = "free", ncol = 7) +
  scale_fill_manual(values = c("dodgerblue","indianred2")) + 
  scale_color_manual(values = c("dodgerblue","indianred2")) + 
  ylab("Frequency") +
  xlab("Condition") +
  theme_classic() +
  ylim(0,NA)+
  theme(strip.background = element_blank(), panel.grid = element_blank()) + 
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) + ggtitle("Spleen Treg FACS") +
  ggpubr::stat_compare_means(label.x = 1.2, label.y.npc = "top", size = 5, vjust = 0.3, label = "p.format") + 
  ggtheme()

#ggsave("figs/spl_facs_treg.svg", width = 36, height = 9.5, units = "cm")
```


```{r}
high_dim2 %>% 
  dplyr::filter(Tissue == "pLN") %>% 
  ggplot(aes(x = Diet, y = value)) + # you can change the x to whatever variable you're interested in
  geom_boxplot(outlier.shape = NA, aes(fill = Diet), alpha = 0.7) + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(width = 0.1, height = 0.1, shape = 21,
              size = 4, aes(fill = Diet)) + # in aes, you can also use shape or fill (for the shapes that allow it)
  facet_wrap(~cluster, scales = "free", ncol = 7) +
  scale_fill_manual(values = c("dodgerblue","indianred2")) + 
  scale_color_manual(values = c("dodgerblue","indianred2")) + 
  ylab("Frequency") +
  xlab("Condition") +
  theme_classic() +
  ylim(0,NA)+
  theme(strip.background = element_blank(), panel.grid = element_blank()) + 
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) + ggtitle("pLN FACS Treg") +
  ggpubr::stat_compare_means(label.x = 1.2, label.y.npc = "top", size = 5, vjust = 0.3, label = "p.format") + 
  ggtheme()

ggsave("figs/pln_facs_treg.svg", width = 36, height = 9.5, units = "cm")
```

```{r}
high_dim2 %>% 
  ggplot(aes(x = Tissue, y = value)) + # you can change the x to whatever variable you're interested in
  geom_boxplot(outlier.shape = NA, aes(fill = Tissue), alpha = 0.7) + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(width = 0.1, height = 0.1, shape = 21,
              size = 4, aes(fill = Tissue)) + # in aes, you can also use shape or fill (for the shapes that allow it)
  facet_wrap(~cluster, scales = "free", ncol = 7) +
  scale_fill_manual(values = c("#56ac00ff","#5fd3bcff","#4a6ca3ff")) + 
  scale_color_manual(values = c("#56ac00ff","#5fd3bcff","#4a6ca3ff")) + 
  ylab("Frequency") +
  xlab("Condition") +
  theme_classic() +
  ylim(0,NA)+
  theme(strip.background = element_blank(), panel.grid = element_blank()) + 
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) + ggtitle("pLN FACS Treg") +
  ggpubr::stat_compare_means(label.x = 1.2, label.y.npc = "top", size = 5, vjust = 0.3, label = "p.format") + 
  ggtheme()

ggsave("figs/facs_treg_tissue.svg", width = 38, height = 9.5, units = "cm")
```

## Treg clones 

```{r}
FeaturePlot(tregs_tfh, reduction = "umap", features = "log_clone_abundance")

FeaturePlot(subset(tregs_tfh, clone_abundance >=1 & clone_abundance < 1056), reduction = "umap", features = "log_clone_abundance", cols = c("lightblue","firebrick")) + theme_classic() + theme(plot.title = element_text(hjust = 0.5)) + ggtitle("Log2 clone abundance")

```


```{r}
metadata_6 <- tregs_tfh@meta.data
rownames(metadata_6) <- colnames(tregs_tfh)

metadata_6 <- metadata_6 %>% mutate(
  clone_nt = paste(cdr3_B_nt,cdr3_A1_nt,cdr3_A2_nt),
  clone_aa = paste("CDR3b",cdr3_B,"CDR3a",cdr3_A1)
)

clone_table_tregs <- metadata_6 %>%
  dplyr::group_by(factor(clone_aa)) %>%
          dplyr::summarize(n = n(), sum = sum()) %>%
    arrange(desc(n))

clone_table_tregs %>% 
  kbl() %>% 
  kable_styling(font_size = 11,
                bootstrap_options = c("striped", "hover", 
                                      "condensed", "responsive")) %>% 
  scroll_box(width = "100%", height = "400px")
```


```{r fig.width=20, fig.height=35}
plot_list <- list()

for(i in 2:15){
 
plot_list[[i]] <-  DimPlot(tregs_tfh, split.by = "Diet",
        cells.highlight = colnames(tregs_tfh)[which(tregs_tfh$clone_aa == pull(clone_table_tregs, `factor(clone_aa)`)[i])],
        sizes.highlight = 2) + NoLegend() + ggtitle(str_replace(string = pull(clone_table_tregs, `factor(clone_aa)`)[i], pattern = "CDR3a", replacement = "\nCDR3a"))
}

cowplot::plot_grid(plotlist = plot_list[2:15], ncol = 2)
```

# Tgd and unconventional

```{r }
DimPlot(funda_merge, group.by = "annotations_fine", label = T)
DimPlot(funda_merge, group.by = "seurat_clusters", label = T)


tgd <- subset(funda_merge, seurat_clusters %in% c(9))

 tgd <- NormalizeData(object = tgd)
  tgd <- ScaleData(tgd, verbose = FALSE)
  tgd <- FindVariableFeatures(tgd, selection.method = "vst", nfeatures = 1000, 
                                       verbose = FALSE)
  tgd <- RunPCA(tgd, npcs = 10, verbose = FALSE)
  tgd <- RunUMAP(tgd, reduction = "pca", dims = 1:10)
  tgd <- FindNeighbors(tgd, dims = 1:10)
  tgd <- FindClusters(tgd, resolution = 0.5)
  DimPlot(tgd, label = T)

  
FeaturePlot(tgd, features = c("Sell", "Cx3cr1"))
FeaturePlot(tgd, features = c("Sell", "Klrg1"))
FeaturePlot(tgd, features = c("Cd8a", "Trdc"))
FeaturePlot(tgd, features = c("Trac", "Tgfb1"))


FeaturePlot(funda_merge, features = c("Trac", "Tgfb1"))


```


```{r }
FeaturePlot(tgd, features = c("Trdc"), pt.size = 12, raster = T,
        raster.dpi = c(1300,1000)) + ggtheme()
ggsave("figv06/tgd_Trdc.svg", width = 13, height = 10, units = "cm")

DimPlot(tgd, cells.highlight = colnames(subset(tgd, cdr3_A1 == "CVVGDRGSALGRLHF")), pt.size = 12, raster = T,
        raster.dpi = c(1200,900)) + ggtheme() + NoLegend()
ggsave("figv06/tgd_iNKT.svg", width = 12, height = 9, units = "cm")
ggsave("figv06/tgd_iNKT.png", width = 12, height = 9, units = "cm")
```


```{r }
DimPlot(tgd, group.by = "Tissue") 
  DimPlot(tgd, group.by = "Diet") 

  grep(rownames(tgd), pattern = "Trg", value=  T)
```


## Markers

```{r}
mrk <- FindAllMarkers(tgd, only.pos = T)

markers_tgd <- FindAllMarkers(tgd, only.pos = T)

markers_tgd
write.csv(markers_tgd, "240806_markers_tgd.csv")
```


```{r  fig.height = 4, fig.width = 5}
funda_merge$tgd <- funda_merge$seurat_clusters %in% c(9)

DimPlot(funda_merge, group.by = "tgd", cols = c("grey78","#a05a2cff"))



DimPlot(tgd, cols = c(
 #"#b9ddf1", 
  #"#b3e0a6",
  "#89b8da", 
  #"#6a9bc3", 
  "#75bc69",
  #"#4878a6" ,
  
  #"#92c0df",
  "#F28E2B" ,
  "#FFBE7D",
  "#9D7660" ,
  "#B07AA1"), pt.size = 12, raster = T,
        raster.dpi = c(900,900)) + ggtheme()
ggsave("fig/tgd_dimplot.svg", width = 8, height = 5.5, units = "cm")

```


```{r}
DimPlot(tgd)

tgd$cdr3_A1
DimPlot(tgd, cells.highlight = colnames(subset(tgd, cdr3_A1 == "CVVGDRGSALGRLHF")))
FeaturePlot(tgd, features = "Il23r")
FeaturePlot(tgd, features = "Ifng")
FeaturePlot(tgd, features = "Tbx21")
FeaturePlot(tgd, features = "Il4")
FeaturePlot(tgd, features = "Ccr6")
FeaturePlot(tgd, features = "Zbtb16")
FeaturePlot(tgd, features = "Ncr1")

  grep(pattern = "Ifng", rownames(tgd), value = T)
```


```{r}
avgexp = AverageExpression(nkt, features = c("Il4", "Ifng", "Gata3", "Tbx21","Il23r","Ccr6"),
                           return.seurat = F, group.by = "sample", 
                          assay = "RNA")

pheatmap(t(avgexp$RNA), main = "",
         scale = "column", cluster_cols = T, cluster_rows = F,
        color=colorRampPalette(c("dodgerblue", "grey95", "indianred2"))(50),
         border_color = "white",
                  fontsize = 12)
```





```{r fig.width = 9, fig.height = 3}
avgexp = AverageExpression(tgd, features = c("Sell","Cd27","Id3","Trdc","Trgv2","Itga4","Klrd1","Sell",
                                             "Gzmb","Gzmk",
                                             "Pdcd1","Eomes","Cd160","Klre1",
                                             
                                             "Tbx21","Ifng","Il4","Cxcr6","Cxcr3","Fasl","Id2","Cd44","Il2ra","Icos","Il7r","Gata3","Zbtb16","Rorc", "Klrg1",
                                             "Ctla4","Nrp1"                                             ),
                           return.seurat = F, group.by = "seurat_clusters", 
                          assay = "RNA")



pheatmap(t(avgexp$RNA)[c(1,4,5,2,3),], main = "",
         scale = "column", cluster_cols = F, cluster_rows = F,
        color=colorRampPalette(c("dodgerblue", "grey95", "indianred2"))(50),
         border_color = "white",
                  fontsize = 12)


pheatmap(t(avgexp$RNA)[c(1,4,5,2,3),], main = "",
         scale = "column", cluster_cols = F, cluster_rows = F,
        color=colorRampPalette(c("dodgerblue", "grey95", "indianred2"))(50),
         border_color = "white",
         fontsize = 12, 
        filename = "figv06/heatmap_tgd_v06.pdf",
         width = 6, height = 2.2,)
```





```{r}

seurat_meta_data <- tgd@meta.data
seurat_meta_data$MouseID2 <- paste(str_replace_all(string = seurat_meta_data$MouseID, 
                                            pattern = "Mouse_", 
                                            replacement = ""), seurat_meta_data$Tissue)
# Create grouped dataframe, calculate the frequencies of clusters
df4 <- seurat_meta_data %>% 
  group_by(MouseID, seurat_clusters) %>% 
  summarise(n = n()) %>% 
  unique() %>% 
  mutate(freq = n / sum(n)) 

# Control - all sums should be one
# df4 %>% group_by(MouseID) %>% summarise(sum = sum(freq))

# As we've lost non-grouping variables, let's join them back
md_to_join <- seurat_meta_data %>% 
  dplyr::select(MouseID, Diet, seurat_clusters, MouseID2, Tissue) %>% 
  unique()

df4  <- left_join(df4, md_to_join)

df4

df4 %>% 
  dplyr::filter(Tissue == "Spleen") %>% 
  ggplot(aes(x = Diet, y = freq*100)) + # you can change the x to whatever variable you're interested in
  geom_boxplot(outlier.shape = NA, aes(fill = Diet), alpha = 0.7) + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(width = 0.1, height = 0.1, shape = 21,
              size = 4, aes(fill = Diet)) + # in aes, you can also use shape or fill (for the shapes that allow it)
  facet_wrap(~seurat_clusters, scales = "free", ncol = 6) +
  scale_fill_manual(values = c("dodgerblue","indianred2")) + 
  scale_color_manual(values = c("dodgerblue","indianred2")) + 
  ylab("Frequency") +
  xlab("Condition") +
  theme_classic() +
  ylim(0,NA)+
  theme(strip.background = element_blank(), panel.grid = element_blank()) + 
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) + ggtitle("Spleen scRNAseq") +
  ggpubr::stat_compare_means(label.x = 1.2, label.y.npc = "top", 
                             method = "t.test", size = 5, vjust = 0.3, label = "p.format") + 
  ggtheme()

  ggsave("figs/tgd_diet_spleen_ttest.png", width = 25, height = 9, units = "cm")

df4 %>% 
  dplyr::filter(Tissue == "pLN") %>% 
  ggplot(aes(x = Diet, y = freq*100)) + # you can change the x to whatever variable you're interested in
  geom_boxplot(outlier.shape = NA, aes(fill = Diet), alpha = 0.7) + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(width = 0.1, height = 0.1, shape = 21,
              size = 4, aes(fill = Diet)) + # in aes, you can also use shape or fill (for the shapes that allow it)
  facet_wrap(~seurat_clusters, scales = "free", ncol = 6) +
  scale_fill_manual(values = c("dodgerblue","indianred2")) + 
  scale_color_manual(values = c("dodgerblue","indianred2")) + 
  ylab("Frequency") +
  xlab("Condition") +
  theme_classic() +
  ylim(0,NA)+
  theme(strip.background = element_blank(), panel.grid = element_blank()) + 
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) + ggtitle("pLN scRNAseq") +
  ggpubr::stat_compare_means(label.x = 1.2, label.y.npc = "top", method = "t.test", 
                             size = 5, vjust = 0.3, label = "p.format") + 
  ggtheme()

  ggsave("figs/tgd_diet_pLN_ttest.png", width = 22, height = 9, units = "cm")

  
  write.csv(df4, "pLN_mw.csv")
```


### Freq from all cells 

```{r}
funda_merge$barcode <- colnames(funda_merge)
df1 <- funda_merge@meta.data %>% dplyr::select(barcode, MouseID, Tissue, Diet)

df1

tgd$barcode <- colnames(tgd)
df2 <- tgd@meta.data %>% dplyr::select(barcode, seurat_clusters)

df3 <- left_join(df1, df2)

df3 <- df3 %>% mutate(seurat_clusters = ifelse(is.na(seurat_clusters),"Other",paste0("Cl. ",seurat_clusters)))


seurat_meta_data <- df3
seurat_meta_data$MouseID2 <- paste(str_replace_all(string = seurat_meta_data$MouseID, 
                                            pattern = "Mouse_", 
                                            replacement = ""), seurat_meta_data$Tissue)
# Create grouped dataframe, calculate the frequencies of clusters
df4 <- seurat_meta_data %>% 
  group_by(MouseID, seurat_clusters) %>% 
  summarise(n = n()) %>% 
  unique() %>% 
  mutate(freq = n / sum(n)) 

# Control - all sums should be one
# df4 %>% group_by(MouseID) %>% summarise(sum = sum(freq))

# As we've lost non-grouping variables, let's join them back
md_to_join <- seurat_meta_data %>% 
  dplyr::select(MouseID, Diet, seurat_clusters, MouseID2, Tissue) %>% 
  unique()

df4  <- left_join(df4, md_to_join)

df4
```


```{r fig.width=12, fig.height=4}
df4 %>% 
  dplyr::filter(Tissue == "Spleen") %>% 
  ggplot(aes(x = Diet, y = freq*100)) + # you can change the x to whatever variable you're interested in
  geom_boxplot(outlier.shape = NA, aes(fill = Diet), alpha = 0.7) + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(width = 0.1, height = 0.1, shape = 21,
              size = 4, aes(fill = Diet)) + # in aes, you can also use shape or fill (for the shapes that allow it)
  facet_wrap(~seurat_clusters, scales = "free", ncol = 6) +
  scale_fill_manual(values = c("dodgerblue","indianred2")) + 
  scale_color_manual(values = c("dodgerblue","indianred2")) + 
  ylab("Frequency") +
  xlab("Condition") +
  theme_classic() +
  ylim(0,NA)+
  theme(strip.background = element_blank(), panel.grid = element_blank()) + 
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) + ggtitle("Spleen scRNAseq") +
  ggpubr::stat_compare_means(label.x = 1.2, label.y.npc = "top", size = 5, vjust = 0.3, label = "p.format") + 
  ggtheme()

  #ggsave("figs/tgd_diet_spleen.svg", width = 25, height = 9, units = "cm")

df4 %>% 
  dplyr::filter(Tissue == "pLN") %>% 
  ggplot(aes(x = Diet, y = freq*100)) + # you can change the x to whatever variable you're interested in
  geom_boxplot(outlier.shape = NA, aes(fill = Diet), alpha = 0.7) + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(width = 0.1, height = 0.1, shape = 21,
              size = 4, aes(fill = Diet)) + # in aes, you can also use shape or fill (for the shapes that allow it)
  facet_wrap(~seurat_clusters, scales = "free", ncol = 6) +
  scale_fill_manual(values = c("dodgerblue","indianred2")) + 
  scale_color_manual(values = c("dodgerblue","indianred2")) + 
  ylab("Frequency") +
  xlab("Condition") +
  theme_classic() +
  ylim(0,NA)+
  theme(strip.background = element_blank(), panel.grid = element_blank()) + 
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) + ggtitle("pLN scRNAseq") +
  ggpubr::stat_compare_means(label.x = 1.2, label.y.npc = "top", size = 5, vjust = 0.3, label = "p.format") + 
  ggtheme()

  #ggsave("figs/tgd_diet_pLN.svg", width = 22, height = 9, units = "cm")

```

## DE NKT1 NKT2

```{r}
DimPlot(tgd, label = T, label.size = 12)
deg_nkt <- FindMarkers(tgd, ident.1 = 1, ident.2 = 2)

deg_nkt

write.csv(deg_nkt, "../2022-10-17_Funda_NOD/tables/deg_nkt.csv")


FeaturePlot(tgd, features = c("Rorc","Xcl1","Ccl5"))
```


### Quantification tissue

```{r}
seurat_meta_data <- tgd@meta.data
# Create grouped dataframe, calculate the frequencies of clusters
df4 <- seurat_meta_data %>% 
  mutate(MouseID_tissue = paste(MouseID, Tissue)) %>% 
  group_by(MouseID_tissue, seurat_clusters) %>% 
  summarise(n = n()) %>% 
  unique() %>% 
  mutate(freq = n / sum(n)) 

# Control - all sums should be one
# df4 %>% group_by(MouseID_tissue) %>% summarise(sum = sum(freq))

# As we've lost non-grouping variables, let's join them back
md_to_join <- seurat_meta_data %>% 
  dplyr::select(MouseID, Diet, Tissue) %>% 
  unique() %>% 
  mutate(MouseID_tissue = paste(MouseID, Tissue))
df4  <- left_join(df4, md_to_join)

# The final plot
df4 %>% 
  ggplot(aes(x = factor(Tissue, labels = c("pLN", "SPL")), y = freq*100)) + # you can change the x to whatever variable you're interested in
  geom_boxplot(outlier.shape = NA, aes(fill = Tissue), alpha = 0.5) +
 geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(shape=21, size = 3, position=position_jitter(width = 0.1, height = 0), 
              aes(fill = Tissue)) + theme_classic() + xlab("Group") +  
  theme_classic() +
  theme(strip.background = element_blank(), panel.grid = element_blank()) + 
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) + ggtitle("Treg scRNAseq by tissue") +
  ggpubr::stat_compare_means(label.x = 1.2, label.y.npc = "top", size = 5, vjust = 0.3, label = "p.format") + 
  ggtheme() +
  scale_color_manual(values = c("#5fd3bcff","#4a6ca3ff")) + 
  scale_fill_manual(values = c("#5fd3bcff","#4a6ca3ff")) +
  facet_wrap(~seurat_clusters, scales = "free", ncol = 6) +
  ylim(0,NA) +
  ylab("Frequency") +
  xlab("Condition") 

  ggsave("figs/tgd_quant_tissue.svg", width = 25, height = 9, units = "cm")

```



# SessionInfo

```{r}
sessionInfo()
```

